---
title: "BEE_Toddler Feasibility"
author: "Jenna Frey"
date: "started 2025-02-07"
output: html_document
---

# FILE SET UP 
## a. Loading libraries and data import
```{r}

# Load packages (list from Elise)
library(openxlsx) # for reading and writing out files
library(table1) # for making tables
library(tidyverse) # for data organization and manipulation
library(ggpubr)  # plots
library(rstatix) # tukeys
library(ltm) # cronbach's alpha
library(FSA) # dunns test
library(patchwork) # for combining graphs
library(ggtext)
library(compareGroups)
# for random forest
library(lubridate)
library(caret)
library(randomForest)
library(e1071)
library(ggsci)
library(themis)
library(this.path)


select <- dplyr::select
rename <- dplyr::rename

# set working directory 
setwd(this.dir())

# main data import
data <- read.csv("1_Data/2025_1217_BEE_toddlerband_feasibility.csv")

# qualitative additions from the band diary 
issues <- read.xlsx("1_Data/QualCode.xlsx")

```

## b. Filtering only BEE participants who participated in the band activity
```{r}

bands <- data %>% 
  filter(bwb_diaryret_T30==1)
  
```

## c. Exploring missingness of the outcome (wear time)
```{r}

# of participants who completed the band visit, who did not return a wrist band?
bands %>% count(bwb_cwret_T30)
#10

# of participants who completed the band visit, who did not return an ankle band?
bands %>% count(bwb_caret_T30)
# 10, 9(==0) and 1(NA)

# of participants who completed the band visit, who did not return both?
bands %>% count(bwb_caret_T30==0 & bwb_cwret_T30==0)
# 6 did not return both bands, 1 had NA responses for both wrist and ankle

# what is the missingness in entire cohort?
all_na <- c(sapply(data, function(x) sum(is.na(x))))
all_vars <- c(names(data))
all_na <- data.frame(all_vars, all_na)

# what is the missingness only among band participants?
band_na <- c(sapply(bands, function(x) sum(is.na(x))))
band_vars <- c(names(bands))
band_na <- data.frame(band_vars, band_na)

# what is the missingness by participant?
bands_indiv_na <- bands %>% 
  rowwise() %>% 
  mutate(var_na=sum(is.na(c_across(bwb_cwret_T30:bwb_pwhrsnitetime7_T30)))) %>% 
  select(S_ID, var_na)

# participant 1023 has missing values for 83 of total columns

# removal of participants who opted out or were missing >75% of data
bands <- bands %>% 
  filter(!S_ID== 1015, # child opted out
         !S_ID== 1023, # missingness shown above
         !S_ID== 1097) # child opted out

# this give a baseline analytical sample size of n=115

```

## d. Setting up dfs
###  1. totals
###  2. separate dfs for each person/band position
###  3. attributes and wear time 
```{r}
# for total, day, and night wear over the week 
wksumwear <- bands %>% 
  select(S_ID, bwb_cwhrsweek_T30, bwb_cwhrsweekday_T30, bwb_cwhrsweeknite_T30, bwb_cahrsweek_T30, bwb_cahrsweekday_T30, bwb_cahrsweeknite_T30, bwb_pwhrsweek_T30, bwb_pwhrsweekday_T30,
bwb_pwhrsweeknite_T30) %>% 
  rename("child_wrist_wktotal" ="bwb_cwhrsweek_T30",
         "child_wrist_daytotal" = "bwb_cwhrsweekday_T30",
         "child_wrist_nighttotal" = "bwb_cwhrsweeknite_T30",
         "child_ankle_wktotal" = "bwb_cahrsweek_T30",
         "child_ankle_daytotal" = "bwb_cahrsweekday_T30",
         "child_ankle_nighttotal" = "bwb_cahrsweeknite_T30",
         "parent_parentwrist_wktotal" = "bwb_pwhrsweek_T30",
         "parent_parentwrist_daytotal" = "bwb_pwhrsweekday_T30", 
         "parent_parentwrist_nighttotal" = "bwb_pwhrsweeknite_T30") %>% 
  pivot_longer(-S_ID, names_to="longname", values_to="hours") %>% 
  separate(longname, c("person", "position", "day"), sep="_")

# separate dfs by person (child, adult) and position (wrist,ankle)
wksum_child_wrist <- wksumwear %>% 
  filter(person == "child", position == "wrist")

wksum_child_ankle <- wksumwear %>% 
  filter(person == "child", position == "ankle")

wksum_adult <- wksumwear %>% 
  filter(person == "parent")


# df of child attributes and weartime totals
attributes_weartimetotals <- bands %>% 
  rename("child_wrist_wktotal" = "bwb_cwhrsweek_T30",
       "child_wrist_daytotal" = "bwb_cwhrsweekday_T30",
       "child_wrist_nighttotal" = "bwb_cwhrsweeknite_T30",
       "child_ankle_wktotal" = "bwb_cahrsweek_T30",
       "child_ankle_daytotal" = "bwb_cahrsweekday_T30",
       "child_ankle_nighttotal" = "bwb_cahrsweeknite_T30",
       "parent_parentwrist_wktotal" = "bwb_pwhrsweek_T30",
       "parent_parentwrist_daytotal" = "bwb_pwhrsweekday_T30", 
       "parent_parentwrist_nighttotal" = "bwb_pwhrsweeknite_T30") %>% 
  select(S_ID, PSI_Mn_T30, PSI_DC_T30, MCDI3_All_T30, ECBQ_dis_T30, ECBQ_EFFCO_T30, ECBQ_NEGAF_T30, ECBQ_SURGE_T30, BIQ_Soc_T30, BIQ_Nov_T30, BIQ_Tot_T30, PAS_AVG_T30, child_wrist_wktotal, child_wrist_daytotal, child_ankle_wktotal, child_ankle_daytotal, child_ankle_nighttotal, parent_parentwrist_wktotal, parent_parentwrist_daytotal, parent_parentwrist_nighttotal) 

```

## e. Exploring missingness (attributes)
```{r}
#wear time and attribute data for ankle
wk_child_ankle_attributes <- merge(wksum_child_ankle, attributes_weartimetotals, by="S_ID", all.x= TRUE) 

#wear time and attribute data for wrist
wk_child_wrist_attributes <- merge(wksum_child_wrist, attributes_weartimetotals, by="S_ID", all.x= TRUE) 

wk_adult_attributes <- merge(wksum_adult, attributes_weartimetotals, by="S_ID", all.x= TRUE)

#attributes only
attributes <- wk_child_ankle_attributes %>% 
  select(S_ID, PSI_Mn_T30:PAS_AVG_T30)

#each participant was in the df 3 times
attributes <- distinct(attributes)

# missingness by attribute
attribute_na <- c(sapply(attributes, function(x) sum(is.na(x))))
attribute_vars <- c(names(attributes))
attribute_na <- data.frame(attribute_vars, attribute_na) %>% 
  rename("attribute"="attribute_na")

# missingness by person
attributes_indiv_na <- attributes %>% 
  rowwise() %>% 
  mutate(var_na=sum(is.na(c_across(PSI_Mn_T30:PAS_AVG_T30)))) %>% 
  select(S_ID, var_na)

  # 12 participants are missing all attributes

```

## f. Evaluating distributions / normality of attributes
### -Histograms
### -Shapiro Wilkes for normality
### -QQ plots
```{r}

###### VISUAL ASSESSMENT

# histograms
attributes_long <- bands %>% 
  select(S_ID, PSI_Mn_T30, PSI_DC_T30, MCDI3_All_T30, ECBQ_dis_T30, ECBQ_EFFCO_T30, ECBQ_NEGAF_T30, ECBQ_SURGE_T30, BIQ_Soc_T30, BIQ_Nov_T30, BIQ_Tot_T30, PAS_AVG_T30,) %>% 
  pivot_longer(!S_ID, names_to="variable", values_to= "value")


attribute_hist <- ggplot(attributes_long, aes(value)) + 
  geom_histogram(fill="cornflowerblue", color= "black", binwidth= function(x) {(max(x)-min(x))/25}) +
  facet_wrap(~ variable, scales= "free") +
  labs( y = "Number of Observations", x= "Value", title= "Distributions of Child Attributes") 

attribute_hist

# qqplot 
attribute_qqplot <- ggqqplot(attributes_long, x= "value", facet.by= "variable", ggtheme= theme_bw(), scales= "free")

attribute_qqplot

###### QUANTITATIVE ASSESSMENT

#shapiro wilkes test  
shapiro_attributes <- apply(attributes_weartimetotals %>% select(-S_ID), 2, shapiro.test)

shapiro_attributes <- do.call(rbind.data.frame, shapiro_attributes)

shapiro_attributes <- shapiro_attributes %>% 
  mutate(normal = ifelse(p.value < 0.05, F, T)) %>%
  select(c(p.value, normal)) 

shapiro_attributes
# normal distributions aside from PSI_DC, MCDI_ALL, ECBQ_Dis

```

## g. creating Table 1: Birthing parent and child demographic characteristics
```{r}

bands %>% count(tc_race)

# checked with the BEE team, and the child missing race data is White, so we will manually change the missing value here: 
bands <- bands %>% 
  mutate(tc_race=ifelse(is.na(tc_race), 0, tc_race))

bands %>% count(tc_race)



# Using same coding criteria as in Hickman et al. (2024) for demographic characteristics of the sample
clean_demo_data <- bands %>%
  select(S_ID:tc_race_other) %>%
  rename("child_age_months_t30" = "TC_AgeMonths_Home_T30", 
         "mat_age_t30" = "PC_matage_home_T30") %>%
  mutate(across(TC_Female:pc_ed, as.character)) %>%
  mutate(TC_Female = recode(TC_Female, 
                            "0" = "M", 
                            "1" = "F")) %>%
  rename("child_sex" = "TC_Female") %>%
  mutate(pc_latino = recode(pc_latino,
                            "0" = "No", 
                            "1" = "MexicanChicano", 
                            "2" = "PuertoRican", 
                            "3" = "Cuban", 
                            "4" = "Other"),
         pc_latino = case_when(pc_latino == "No" ~ "No", 
                               .default = "Yes"),
         tc_latino = recode(tc_latino, 
                            "0" = "No", 
                            "1" = "MexicanChicano", 
                            "2" = "PuertoRican", 
                            "3" = "Cuban", 
                            "4" = "Other"),
         tc_latino = case_when(tc_latino == "No" ~ "No", 
                               .default = "Yes"),
         pc_race = recode(pc_race, "0" = "White", 
                          "1" = "Black", 
                          "2" = "IndigenousAmerican", 
                          "3" = "AsianIndian", 
                          "4" = "Chinese", 
                          "5" = "Japanese", 
                          "6" = "Korean", 
                          "7" = "Vietnamese", 
                          "8" = "OtherAsian", 
                          "9" = "Other"),
         pc_biracial = recode(pc_biracial, "0" = "No", 
                              "1" = "Yes"),
         pc_addrace = recode(pc_addrace, "0" = "White", 
                             "1" = "Black", 
                             "2" = "IndigenousAmerican", 
                             "3" = "AsianIndian", 
                             "4" = "Chinese", 
                             "5" = "Japanese", 
                             "6" = "Korean", 
                             "7" = "Vietnamese", 
                             "8" = "OtherAsian", 
                             "9" = "Other"),
         pc_race_black = recode(pc_race_black, 
                                "0" = "No", 
                                "1" = "Yes"),
         pc_ed = recode(pc_ed, 
                        "11" = "High School, GED, or less",
                        "12" = "High School, GED, or less", 
                        "13" = "High School, GED, or less", 
                        "14" = "Associate Degree", 
                        "15" = "Associate Degree", 
                        "16" = "Four-Year Degree", 
                        "17" = "Four-Year Degree", 
                        "18" = "Master's Degree", 
                        "19" = "Professional Degree or PhD"),
         tc_race = recode(tc_race, 
                          "0" = "White", 
                          "1" = "Black", 
                          "2" = "IndigenousAmerican", 
                          "3" = "AsianIndian", 
                          "4" = "Chinese", 
                          "5" = "Japanese", 
                          "6" = "Korean", 
                          "7" = "Vietnamese", 
                          "8" = "OtherAsian", 
                          "9" = "Other"),
         tc_biracial = recode(tc_biracial, 
                              "0" = "No", 
                              "1" = "Yes"),
         tc_addrace = recode(tc_addrace, 
                             "0" = "White", 
                             "1" = "Black", 
                             "2" = "IndigenousAmerican", 
                             "3" = "AsianIndian", 
                             "4" = "Chinese", 
                             "5" = "Japanese", 
                             "6" = "Korean", 
                             "7" = "Vietnamese", 
                             "8" = "OtherAsian", 
                             "9" = "Other"),
         tc_race_black = recode(tc_race_black, 
                                "0" = "No", 
                                "1" = "Yes")
         ) %>%
  select(-c(pc_race_black, tc_race_black)) %>%
  mutate(pc_latino_hispanic = case_when(pc_latino == "Yes" ~ "Yes",
                                        pc_race_other == "Hispanic" ~ "Yes",
                                        pc_race_other == "Mexican" ~ "Yes",
                                        pc_race_other == "Latina" ~ "Yes",
                                        .default = "No"),
         pc_multiracial_cleaned = case_when(pc_race_other == "biracial" ~ "Yes",
                                            pc_race_other == "multi" ~ "Yes",
                                            pc_biracial == "Yes" ~ "Yes",
                                            .default = "No"),
         pc_race_cleaned = case_when(pc_multiracial_cleaned == "Yes" ~ "Biracial/Multiracial",
                                     pc_race == "White" ~ "White",
                                     pc_race == "Black" ~ "Black",
                                     pc_race == "IndigenousAmerican" ~ "Other", 
                                     pc_race == "AsianIndian" ~ "Other",
                                     pc_race == "Chinese" ~ "Other",
                                     pc_race == "Japanese" ~ "Other", 
                                     pc_race == "Korean" ~ "Other",
                                     pc_race == "Vietnamese" ~ "Other",
                                     pc_race =="OtherAsian" ~ "Other",
                                     pc_race == "Other" ~ "Other")) %>%
  mutate(tc_latino_hispanic = case_when(tc_latino == "Yes" ~ "Yes",
                                        tc_race_other == "Hispanic" ~ "Yes",
                                        tc_race_other == "Mexican" ~ "Yes",
                                        tc_race_other == "Latina" ~ "Yes",
                                        .default = "No"),
         tc_multiracial_cleaned = case_when(tc_race_other == "biracial" ~ "Yes",
                                            tc_race_other == "multi" ~ "Yes",
                                            tc_biracial == "Yes" ~ "Yes",
                                            .default = "No"),
         tc_race_cleaned = case_when(tc_multiracial_cleaned == "Yes" ~ "Biracial/Multiracial",
                                     tc_race == "White" ~ "White",
                                     tc_race == "Black" ~ "Black",
                                     tc_race == "Chinese" ~ "Other",
                                     tc_race == "Other" ~ "Other")) %>%
  select(-c(pc_latino:pc_addrace, pc_race_other, last_pc_ed:pc_race_other, pc_multiracial_cleaned, tc_latino:tc_race_other, tc_multiracial_cleaned))

#from EH 7/18
# Split demographics table into child and mother tables, rename columns so they match, and create grouping column to indicate whether data go with mother or child
clean_demo_data_tc <- clean_demo_data %>%
  select(c(S_ID, child_age_months_t30, child_sex, tc_latino_hispanic, tc_race_cleaned)) %>%
  rename("age" = "child_age_months_t30", "sex" = "child_sex", "latino_hispanic" = "tc_latino_hispanic", "race" = "tc_race_cleaned") %>%
  mutate(group = "Child") %>%
  mutate(education = NA)

clean_demo_data_pc <- clean_demo_data %>%
  select(c(S_ID, mat_age_t30, pc_ed, pc_latino_hispanic, pc_race_cleaned)) %>%
  rename("age" = "mat_age_t30", "latino_hispanic" = "pc_latino_hispanic", "race" = "pc_race_cleaned", "education" = "pc_ed") %>%
  mutate(sex = "F", group = "Parent")

# Merge together and level factors
clean_demo_data_long <- bind_rows(clean_demo_data_pc, clean_demo_data_tc) %>%
  mutate(race = factor(race, levels = c("White", "Black", "Biracial/Multiracial", "Other")),
         latino_hispanic = factor(latino_hispanic, levels = c("No", "Yes")),
         education = factor(education, levels=c("High School, GED, or less", "Associate Degree", 
                                                "Four-Year Degree", "Master's Degree","Professional Degree or PhD")),
         sex = factor(sex, levels = c("F", "M")),
         group = factor(group, levels = c("Parent", "Child")))

# Add labels
label(clean_demo_data_long$age) <- "Age"
label(clean_demo_data_long$education) <- "Education"
label(clean_demo_data_long$latino_hispanic) <- "Latino or Hispanic"
label(clean_demo_data_long$race) <- "Race"
label(clean_demo_data_long$sex) <- "Sex"

# Create function for custom table so that Mean (SD) is shown for continuous variables
my.render.cont <- function(x) {
  with(stats.apply.rounding(stats.default(x), digits=2), c("",
                                                           "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

# Make table of demographics
table1(~ sex + age + race + latino_hispanic + education | group,
       data = clean_demo_data_long, 
       render.continuous = my.render.cont, 
       overall = FALSE)

```



## first half of this code chunk written by allison spring to compare those who are in our analytical sample to those who were not 
```{r}

data %>% count(Complete_Home_T30)

#These are the S_IDs of the 115 
my_list<-bands$S_ID

included_subjects <- data[data$S_ID %in% my_list, ] %>%
  dplyr::filter(Complete_Home_T30==1)


Removed_subjects <- data[!data$S_ID %in% my_list, ] %>%
  filter(Complete_Home_T30==1)


included_subjects <- lapply(included_subjects, as.numeric)
Removed_subjects  <- lapply(Removed_subjects, as.numeric)

Removed_subjects <- as.data.frame(Removed_subjects)
included_subjects <- as.data.frame(included_subjects)

# Identify numeric columns (skip non-numeric)
numeric_cols <- sapply(Removed_subjects, is.numeric)

# Subset the data frame to include only numeric columns
df_numeric <- Removed_subjects[, numeric_cols]

# Compute the column means for only numeric columns
averages <- colMeans(df_numeric, na.rm = TRUE)

# Print the result
print(averages)

# Compute the average of each column in both datasets
averages_included_subjects <- colMeans(included_subjects, na.rm = TRUE)
averages_Removed_subjects <- colMeans(Removed_subjects, na.rm = TRUE)

# Combine the averages into a data frame for better visualization
averages_combined <- data.frame(
  averages_included_subjects = averages_included_subjects,
  averages_Removed_subjects = averages_Removed_subjects
)

# Print the averages
print(averages_combined)


# jcf
Removed_subjects$in_sample <- "0"
included_subjects$in_sample <- "1"

all_subjects <- rbind(Removed_subjects, included_subjects)
all_subjects <- all_subjects %>% 
  mutate(across(TC_Female:tc_race_other, as.factor))

# comparison using compareGroups
table1_comparison <- compareGroups(in_sample ~ . - S_ID, data= all_subjects, method=4)
  # method = 4 does a shapiro wilks test to determine normality of each var

table1_comparison

createTable(table1_comparison)

hist(all_subjects$income2needs_T30)
wilcox.test(income2needs_T30 ~in_sample, data=all_subjects, alternative="two.sided")

```





# RQ1: HOW COMPLIANT WERE TODDLERS WITH WEARING THE SILICONE WRISTBAND AND ANKLE BAND (AS DOCUMENTED BY THE PARENT-COMPLETED DAILY DIARIES)?

## a. Histograms, univariate distributions, and summary statistics (mean, median, min, max, std) for wear time by person (child, parent) and band position (wrist, ankle)
```{r}
###### VISUAL ASSESSMENT: histograms

# CHILD WRIST
#day
hist(bands$bwb_cwhrsweekday_T30)

#night
hist(bands$bwb_cwhrsweeknite_T30)

#total 
hist(bands$bwb_cwhrsweek_T30)

# CHILD ANKLE 
#day
hist(bands$bwb_cahrsweekday_T30)

#night
hist(bands$bwb_cahrsweeknite_T30)

#total 
hist(bands$bwb_cahrsweek_T30)

# PARENT WRIST
#day
hist(bands$bwb_pwhrsweekday_T30)

#night
hist(bands$bwb_pwhrsweeknite_T30)

#total 
hist(bands$bwb_pwhrsweek_T30)

```

```{r}
##### QUANTITATIVE ASSESSMENT

# summary statistics for each person (child, parent), band position (wrist, ankle), and time (day total [max 84], night total [max 84], week total [max 168])
sumstat_totalweekwear <- wksumwear %>% 
  group_by(day, person, position) %>% 
  summarise(median=median(hours, na.rm=TRUE), min= min(hours, na.rm=TRUE), max=max(hours, na.rm=TRUE), mean=mean(hours, na.rm= TRUE), std= sd(hours, na.rm=TRUE))

sumstat_totalweekwear

## comparison of wrist and ankle wear time overall and within time of day (day, night) strata ##

# creating a new df 
wksumsep <- wksumwear %>% 
  filter(!(day=="wktotal"),                                              # only day and night wear
         !(person=="parent")) %>%                                        # only children
  mutate(day = case_when(day == "daytotal" ~ "Day",                      # cleaning of var names for figure 1 later
                        day == "nighttotal" ~ "Night")) %>% 
  mutate(position = case_when(position == "wrist" ~ "Child Wrist",
                        position == "ankle" ~ "Child Ankle",
                        position == "parentwrist" ~ "Parent Wrist"))

## OVERALL COMPARISON 

# unpaired comparison of hours worn between the wrist and ankle band overall (day and night together)
wilcox.test(hours ~ position, data=wksumsep)

# paired comparison of hours worn between the wrist and ankle band overall (day and night together)
wilcox.test(x = wksumsep$hours[wksumsep$position == "Child Wrist"],  
            y = wksumsep$hours[wksumsep$position == "Child Ankle"],  
            paired = TRUE)


## WITHIN TIME OF DAY STRATA 

# DAYTIME

# new df of only day
wksumsep_day <- wksumsep %>% 
  filter(day=="Day")

# unpaired comparison of hours worn between the wrist and ankle band during the day
wilcox.test(hours ~ position, data=wksumsep_day)

# paired comparison of hours worn between the wrist and ankle band during the day
wilcox.test(x = wksumsep_day$hours[wksumsep_day$position == "Child Wrist"],  
            y = wksumsep_day$hours[wksumsep_day$position == "Child Ankle"],  
            paired = TRUE)



# NIGHTTIME

# new df of only night 
wksumsep_night <- wksumsep %>% 
  filter(day=="Night")

# unpaired comparison of hours worn between the wrist and ankle band at night
wilcox.test(hours ~ position, data=wksumsep_night)

# paired comparison of hours worn between the wrist and ankle band at night 
wilcox.test(x = wksumsep_night$hours[wksumsep_night$position == "Child Wrist"],  
            y = wksumsep_night$hours[wksumsep_night$position == "Child Ankle"],  
            paired = TRUE)





# Within wrist band, is wear time different by time of day?
wksumsep_child_wrist <- wksum_child_wrist %>% 
  filter(!day=="wktotal")

# unpaired comparison of hours worn between the day and night for the wrist band
wilcox.test(hours ~ day, data=wksumsep_child_wrist)

# paired comparison of hours worn between the day and night for the wrist band
wilcox.test(x = wksumsep_child_wrist$hours[wksumsep_child_wrist$day == "daytotal"],  
            y = wksumsep_child_wrist$hours[wksumsep_child_wrist$day == "nighttotal"],  
            paired = TRUE)


# Within ankle band, is wear time different by time of day?
wksumsep_child_ankle <- wksum_child_ankle %>% 
  filter(!day=="wktotal")

# paired comparison of hours worn between the day and night for the ankle band
wilcox.test(hours ~ day, data=wksumsep_child_ankle)

# paired comparison of hours worn between the day and night for the ankle band
wilcox.test(x = wksumsep_child_ankle$hours[wksumsep_child_ankle$day == "daytotal"],  
            y = wksumsep_child_ankle$hours[wksumsep_child_ankle$day == "nighttotal"],  
            paired = TRUE)

```
### Primary focus is on DAYTIME WEAR ONLY
## -new dataframes of only day time wear by each day (1-7) 
## -summary statistics of day time wear
## -correlation between number of days worn and total week wear time
```{r}
# new long df of day time wear by each day with cleaned variable names
daywear <- bands %>% 
  select(S_ID, contains("daytime")) %>% 
  rename("child_wrist_1" = "bwb_cwhrsdaytime1_T30",                      
         "child_ankle_1" = "bwb_cahrsdaytime1_T30", 
         "parent_wrist_1"= "bwb_pwhrsdaytime1_T30",
         "child_wrist_2" = "bwb_cwhrsdaytime2_T30",
         "child_ankle_2" = "bwb_cahrsdaytime2_T30",
         "parent_wrist_2" = "bwb_pwhrsdaytime2_T30",
         "child_wrist_3" = "bwb_cwhrsdaytime3_T30", 
         "child_ankle_3" = "bwb_cahrsdaytime3_T30", 
         "parent_wrist_3" = "bwb_pwhrsdaytime3_T30",
         "child_wrist_4" = "bwb_cwhrsdaytime4_T30", 
         "child_ankle_4" = "bwb_cahrsdaytime4_T30", 
         "parent_wrist_4" = "bwb_pwhrsdaytime4_T30",
         "child_wrist_5" = "bwb_cwhrsdaytime5_T30", 
         "child_ankle_5" = "bwb_cahrsdaytime5_T30", 
         "parent_wrist_5" = "bwb_pwhrsdaytime5_T30",
         "child_wrist_6" = "bwb_cwhrsdaytime6_T30", 
         "child_ankle_6" = "bwb_cahrsdaytime6_T30", 
         "parent_wrist_6" = "bwb_pwhrsdaytime6_T30",
         "child_wrist_7" = "bwb_cwhrsdaytime7_T30", 
         "child_ankle_7" ="bwb_cahrsdaytime7_T30", 
         "parent_wrist_7" = "bwb_pwhrsdaytime7_T30") %>% 
  pivot_longer(-S_ID, names_to="longname", values_to="hours") %>% 
  separate(longname, c("person", "position", "day"), sep="_")

# summary statistics for daywear data (day by day)
    ## median, minimum, maximum, mean, and standard deviation
sumstat_day <- daywear %>% 
  group_by(day, person, position) %>% 
  summarise(median=median(hours, na.rm=TRUE), min= min(hours, na.rm=TRUE), max=max(hours, na.rm=TRUE), mean=mean(hours, na.rm= TRUE), std= sd(hours, na.rm=TRUE))



# counts of the days by band/person with missing values
daywear %>% group_by(day, person, position) %>% count(is.na(hours))
#day 1,2,3,6 no missing 
#day 4 missing reported hours for 1 child wrist, 1 child ankle, and 1 parent wrist band (all same family--1153)
#day 5 missing reported hours for 2 child wrist, 2 child ankle, and 2 parent wrist band (all same families-- 1153, 1059)
#day 7 missing reported hours for 3 child wrist, 3 child ankle, and 3 parent wrist band (all same families-- 1153, 1116, 1029)


# wide df of day time wear for each participant by day with counts of how many times they fall in each of the following categories across the week (max 7)
  # 0-1 hours    = low 
  # 2-10 hours   = mid
  # 11-12 hours  = high
daywear_short <- daywear %>% 
  pivot_wider(names_from= day, values_from= hours) %>% 
   rename(day1="1", 
         day2="2",
         day3="3",
         day4="4",
         day5="5",
         day6="6", 
         day7="7") %>%  
  mutate(n.na=rowSums(is.na(.)), 
         n.0_1=rowSums(select(.,c(day1:day7)) < 2, na.rm=TRUE),
         n.2_10=rowSums((select(.,c(day1:day7))) < 11 & (select(.,c(day1:day7))) > 1, na.rm=TRUE), 
         n.11_12=rowSums(select(.,c(day1:day7)) > 10, na.rm=TRUE), 
         n.2_12=rowSums(select(.,c(day1:day7)) > 1, na.rm=TRUE)) %>% 
  rowwise() %>% 
  mutate(n.total=sum(c_across(c(n.na:n.11_12)))) %>% #check
  ungroup()

# creating a new binary variable of worn/not worn for the day where
## worn= 2 or more hours a day 
## not worn= less than 2 hours a day 
daywear_short_binary <- daywear_short %>% 
  mutate(day1b=case_when(day1== NA ~ NA, 
                         day1 < 2 ~ 0,
                         day1 >= 2 ~ 1),
         day2b=case_when(day2== NA ~ NA, 
                         day2 < 2 ~ 0,
                         day2 >= 2 ~ 1), 
         day3b=case_when(day3== NA ~ NA, 
                         day3 < 2 ~ 0,
                         day3 >= 2 ~ 1), 
         day4b=case_when(day4== NA ~ NA, 
                         day4 < 2 ~ 0,
                         day4 >= 2 ~ 1), 
         day5b=case_when(day5== NA ~ NA, 
                         day5 < 2 ~ 0,
                         day5 >= 2 ~ 1), 
         day6b=case_when(day6== NA ~ NA, 
                         day6 < 2 ~ 0,
                         day6 >= 2 ~ 1), 
         day7b=case_when(day7== NA ~ NA, 
                         day7 < 2 ~ 0,
                         day7 >= 2 ~ 1)) %>% 
  mutate(n.daysworn=rowSums(select(.,c(day1b:day7b)), na.rm=TRUE))

## 8 hour threshold 
daywear_short_binary <- daywear_short_binary %>% 
  mutate(day1_8=case_when(day1 == NA ~ NA, 
                          day1 <  8 ~ 0,
                          day1 >= 8 ~ 1),
         day2_8=case_when(day2 == NA ~ NA, 
                          day2 <  8 ~ 0,
                          day2 >= 8 ~ 1), 
         day3_8=case_when(day3 == NA ~ NA, 
                          day3 <  8 ~ 0,
                          day3 >= 8 ~ 1), 
         day4_8=case_when(day4 == NA ~ NA, 
                          day4 <  8 ~ 0,
                          day4 >= 8 ~ 1), 
         day5_8=case_when(day5 == NA ~ NA, 
                          day5 <  8 ~ 0,
                          day5 >= 8 ~ 1), 
         day6_8=case_when(day6 == NA ~ NA, 
                          day6 <  8 ~ 0,
                          day6 >= 8 ~ 1), 
         day7_8=case_when(day7 == NA ~ NA, 
                         day7 <  8 ~ 0,
                         day7 >= 8 ~ 1)) %>% 
  mutate(n.daysworn_8=rowSums(select(.,c(day1_8:day7_8)), na.rm=TRUE))








# selecting only the daytime weekly wear time total variable for ankle and wrist from "bands" df in order to merge it with the "daywear_short_binary" df
bands_child_daytotals <- bands %>% 
  select(S_ID, bwb_cwhrsweekday_T30, bwb_cahrsweekday_T30) %>% 
  rename("child_wrist_daytotal" = "bwb_cwhrsweekday_T30", 
         "child_ankle_daytotal" = "bwb_cahrsweekday_T30")

# merge
daywear_short_binary <- merge(daywear_short_binary, bands_child_daytotals, all=TRUE)

# visual assessment of normality 
## non normal, will proceed with a spearman test for correlations
hist(daywear_short_binary$n.daysworn)
hist(daywear_short_binary$child_wrist_daytotal)
hist(daywear_short_binary$child_ankle_daytotal)

daywear_short_binary_ankle <- daywear_short_binary %>% 
  filter(person=="child", 
         position=="ankle")

daywear_short_binary_wrist <- daywear_short_binary %>% 
  filter(person=="child", 
         position=="wrist")

# is the number of days worn in a week correlated with the total hours worn over the week for the ankle band?
cor.test(daywear_short_binary_ankle$n.daysworn, daywear_short_binary_ankle$child_ankle_daytotal, method="spearman")

# plot
daywear_corr_ankle <- ggplot(daywear_short_binary_ankle, aes(x=n.daysworn, y=child_ankle_daytotal, color=position)) + 
  geom_point(alpha= 0.7, size=3) +
  scale_color_manual(values = ("ankle" = "cornflowerblue")) +
  stat_cor(method= "spearman", label.x = 1, label.y = 80, color= "black") +
  geom_point()

daywear_corr_ankle

# is the number of days worn in a week correlated with the total hours worn over the week for the wrist band?
cor.test(daywear_short_binary_wrist$n.daysworn, daywear_short_binary_wrist$child_wrist_daytotal, method="spearman")

# plot
daywear_corr_wrist <- ggplot(daywear_short_binary_wrist, aes(x=n.daysworn, y=child_wrist_daytotal, color=position)) + 
  geom_point(alpha= 0.7, size=3) +
  scale_color_manual(values = ("wrist" = "orange")) +
  stat_cor(method= "spearman", label.x = 1, label.y = 80, color= "black") 

daywear_corr_wrist

# combined plot of wrist/ankle
corr_by_day <- daywear_corr_ankle + daywear_corr_wrist +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 18, face = "bold"))
  
corr_by_day

# jpeg("2_OutputGraphs/corr_by_day.jpeg",
#     width = 8, height = 4, units="in", res=1200)
# corr_by_day
# invisible(dev.off())



# ankle only min and max for total hours across the week
daywear_short_sum_ankle <- daywear_short_binary_ankle %>% 
  group_by(n.daysworn) %>% 
  summarise(min_ankle=min(child_ankle_daytotal), max_ankle=max(child_ankle_daytotal))

# wrist only min and max for total hours across the week
daywear_short_sum_wrist <- daywear_short_binary_wrist %>% 
  group_by(n.daysworn) %>% 
  summarise(min_wrist=min(child_wrist_daytotal), max_wrist=max(child_wrist_daytotal))

# number of participants who wore the band for 0-7 days for the wrist and ankle bands
daywear_short_binary_wrist %>% count(n.daysworn)
daywear_short_binary_ankle %>% count(n.daysworn)

# combined df of min and max of total hours worn in the week by the number of days worn
daywear_short_sum <- merge(daywear_short_sum_ankle, daywear_short_sum_wrist)

daywear_short_sum




# ankle only min and max for total hours across the week
daywear_short_sum_ankle8 <- daywear_short_binary_ankle %>% 
  group_by(n.daysworn_8) %>% 
  summarise(min_ankle=min(child_ankle_daytotal), max_ankle=max(child_ankle_daytotal))

# wrist only min and max for total hours across the week
daywear_short_sum_wrist8 <- daywear_short_binary_wrist %>% 
  group_by(n.daysworn_8) %>% 
  summarise(min_wrist=min(child_wrist_daytotal), max_wrist=max(child_wrist_daytotal))

# number of participants who wore the band for 0-7 days for the wrist and ankle bands
daywear_short_binary_wrist %>% count(n.daysworn_8)
daywear_short_binary_ankle %>% count(n.daysworn_8)

# combined df of min and max of total hours worn in the week by the number of days worn
daywear_short_sum_8 <- merge(daywear_short_sum_ankle8, daywear_short_sum_wrist8)

daywear_short_sum_8


```

## b. reliability across the week (cronbach's alpha)
```{r}
#dfs by ankle and wrist for day 1 through 7 
daywear_days1_7_ankle <- daywear_short %>% 
  filter(position=="ankle") %>% 
  select(day1:day7) 
 
daywear_days1_7_wrist <- daywear_short %>% 
  filter(position=="wrist", 
         person=="child") %>% 
  select(day1:day7) 
  
#dfs by ankle and wrist for day 2 through 7
daywear_days2_7_ankle <- daywear_days1_7_ankle %>% 
  select(-day1) 

daywear_days2_7_wrist <- daywear_days1_7_wrist %>% 
  select(-day1)

#ankle reliability test (cronbach's alpha)

#days 1 through 7
cronbach.alpha(daywear_days1_7_ankle, CI=TRUE, na.rm=TRUE)
#days 2 through 7
cronbach.alpha(daywear_days2_7_ankle, CI=TRUE, na.rm=TRUE)

#wrist reliability test (cronbach's alpha)

#days 1 through 7
cronbach.alpha(daywear_days1_7_wrist, CI=TRUE, na.rm=TRUE)
#days 2 through 7
cronbach.alpha(daywear_days2_7_wrist, CI=TRUE, na.rm=TRUE)

```
## c. wear time distributions
### -justification of split at 72 hrs (6 days) based on the histogram
### -creating new binary variable at 72 hours
### -Spearman correlations and scatter plots for
####  1. child wrist vs. child ankle 
####  2. child wrist vs. parent wrist 
####  3. child ankle vs. parent wrist
```{r}

# combining ankle and wrist dfs that have attribute and day, night, and total hours for ONLY the child
wk_child_attributes <- bind_rows(wk_child_ankle_attributes, wk_child_wrist_attributes)

# combining ankle and wrist dfs that have attribute and day, night, and total hours for child and parent
wk_all_attributes <- bind_rows(wk_child_ankle_attributes, wk_child_wrist_attributes, wk_adult_attributes)

# selecting only daytime totals (focus of this analysis) and the child attributes
wk_child_day <- wk_child_attributes %>% 
  filter(person == "child" & day== "daytotal") %>% 
  select(-day,
         -person, 
         -contains("total"))

hist(wk_child_day$hours, breaks=20) # shows a split at 72 hours which is 6 days of wear

# new binary variable at 72 hours
wk_child_day <- wk_child_day %>% 
  mutate(cat_hours = case_when(hours >= 72 ~ "72 or more hours", 
                               hours < 72 ~ "< 72 hours"))

# check
wk_child_day %>% count(cat_hours)

# reorder variables in df
wk_child_day <- wk_child_day %>% 
  relocate(cat_hours, .before=PSI_Mn_T30)

# long df to plot 
long_wk_child_day <- wk_child_day %>% 
  select(-contains("total")) %>% 
  pivot_longer(!c(S_ID, position, cat_hours, hours), names_to = "attribute", values_to = "value")


# scatterplot of day time week long wear (y) for each attribute (scale on the x) by child wrist (orange) and ankle band (blue)
attribute_scat <- ggplot(long_wk_child_day, aes(x=value, y=hours, color=position, alpha= 0.5)) + 
  scale_color_manual(values = c("wrist" = "orange", 
                               "ankle" = "cornflowerblue")) + 
  geom_point() + 
  facet_wrap(~attribute, scales= "free")

attribute_scat

# df of adult hours with attributes to assess whether child wear time and parent wear time are correlated
wk_adult_day <- wk_all_attributes %>% 
  filter(person == "parent" & day== "daytotal") %>% 
  select(-day,
         -person)

# df of adult hours and SID only
wk_adult_day_hours <- wk_adult_day %>% 
  select(S_ID, hours) %>% 
  rename("adult_hours" = "hours")

# separating wrist and ankle to rename/create a new dataframe

# ankle only 
wk_child_day_ankle <- wk_child_day %>% 
  filter(position == "ankle")

wk_child_day_ankle_hours <- wk_child_day_ankle %>% 
  select(S_ID, hours, cat_hours) %>% 
  rename("ankle_hours" = "hours", 
         "ankle_cat_hours" = "cat_hours")

#wrist only
wk_child_day_wrist <- wk_child_day %>% 
  filter(position == "wrist")

wk_child_day_wrist_hours <- wk_child_day_wrist %>% 
  select(S_ID, hours, cat_hours) %>% 
  rename("wrist_hours" = "hours",
         "wrist_cat_hours" = "cat_hours")

wk_child_day_hours <- merge(wk_child_day_ankle_hours, wk_child_day_wrist_hours, all= TRUE)

wk_all_day_hours <- merge(wk_child_day_hours, wk_adult_day_hours, all= TRUE)



# what is the average difference across participants for ankle vs. wrist bands?
wk_child_day_hours$difference <- wk_child_day_hours$ankle_hours-wk_child_day_hours$wrist_hours 

mean(wk_child_day_hours$difference)

# **** paired wilcox sign rank test is presented after demo_weartime df is created ****




# child wrist to ankle spearman correlation
cor.test(wk_child_day_hours$ankle_hours, wk_child_day_hours$wrist_hours, method="spearman") 

# parent wrist to child wrist spearman correlation 
cor.test(wk_all_day_hours$adult_hours, wk_all_day_hours$wrist_hours, method="spearman") 

# parent wrist to child ankle spearman correlation
cor.test(wk_all_day_hours$adult_hours, wk_all_day_hours$ankle_hours, method="spearman") 



#CORRESPONDING PLOTS 

#wrist by ankle wear time
anklebywrist_scatter <- ggplot(wk_child_day_hours, aes(x= wrist_hours, y=ankle_hours)) + 
  geom_point(size= 3) + 
  stat_cor(method= "spearman", label.x = 3, label.y = 30) + 
  labs(title= "Child Band Daytime Wear")  

anklebywrist_scatter

#wrist by parent wear time 
wristbyparent_scatter <- ggplot(wk_all_day_hours, aes(x= wrist_hours, y=adult_hours)) + 
  geom_point(size= 3, color="orange") + 
  stat_cor(method= "spearman", label.x = 3, label.y = 30) + 
  labs(title= "Child wrist by Adult wrist Band Daytime Wear")  

wristbyparent_scatter

#ankle by parent wear time
anklebyparent_scatter <- ggplot(wk_all_day_hours, aes(x= ankle_hours, y=adult_hours)) + 
  geom_point(size= 3, color="cornflowerblue") + 
  stat_cor(method= "spearman", label.x = 3, label.y = 30) + 
  labs(title= "Child ankle by Adult wrist Band Daytime Wear")  

anklebyparent_scatter

```

## d. Low, moderate, and high wear time categorization by median wear time 
### Table 2
```{r}

median_cat_weartime <-  daywear %>% 
  group_by(S_ID, position, person) %>% 
  summarise(median= median(hours, na.rm=TRUE)) %>%  # median wear time across the 7 days 
  # for missing individual days, missing days were removed when calculating the median wear time
      # 3 participants were missing 1 day 
      # 1 participant was missing 3 days 
  mutate(hourscat= case_when(median < 2 ~ "0-1",
                            median < 11 ~ "2-10",
                            median > 10 ~ "11-12"))
# check
median_cat_weartime %>% 
  group_by(position, person) %>% 
  count(hourscat) 

# separating by position to make the matrix
# ankle
median_cat_weartime_ankle <-  median_cat_weartime %>% 
  filter(!person=="parent", 
         position=="ankle") %>% 
  rename(ankle_hourscat= hourscat) %>% 
  ungroup(position) %>% 
  select(S_ID, ankle_hourscat)

# wrist
median_cat_weartime_wrist <-  median_cat_weartime %>% 
  filter(!person=="parent", 
         position=="wrist") %>% 
  rename(wrist_hourscat= hourscat) %>%
  ungroup(position) %>% 
  select(S_ID, wrist_hourscat)


median_cat_weartime_matrix <- merge(median_cat_weartime_ankle, median_cat_weartime_wrist, all= TRUE) %>% 
  mutate(combined_med = case_when(ankle_hourscat == "11-12" & wrist_hourscat == "11-12" ~ "high ankle & high wrist", 
                                ankle_hourscat == "11-12" & wrist_hourscat == "2-10" ~ "high ankle & med wrist",
                                ankle_hourscat == "11-12" & wrist_hourscat == "0-1" ~ "high ankle & no/low wrist",
                                ankle_hourscat == "2-10" & wrist_hourscat == "11-12" ~ "med ankle & high wrist",
                                ankle_hourscat == "2-10" & wrist_hourscat == "2-10" ~ "med ankle & med wrist", 
                                ankle_hourscat == "2-10" & wrist_hourscat == "0-1" ~ "med ankle & no/low wrist",
                                ankle_hourscat == "0-1" & wrist_hourscat == "11-12" ~ "no/low ankle & high wrist",
                                ankle_hourscat == "0-1" & wrist_hourscat == "2-10" ~ "no/low ankle & med wrist", 
                                ankle_hourscat == "0-1" & wrist_hourscat == "0-1" ~ "no/low ankle & no/low wrist"))
median_cat_weartime_matrix %>% 
  count(combined_med)

```

## e. Low, moderate, and high wear time categorizations by total week wear
### Table 2
```{r}

# new categorical variable for consistency with the <6 days or ≥6 days of wear time over the week
totalhours_cat <- wk_child_day %>% 
  select(S_ID:hours) %>% 
  mutate(threecat_hours= case_when(hours < 12 ~ "low", #less than one full day
                              hours >= 12 & hours < 72 ~ "med", #mid 
                              hours >= 72 ~ "high")) #six or more days

totalhours_cat$position_cathours <- paste(totalhours_cat$position, totalhours_cat$threecat_hours)

# now need to position the dfs next to each other rather than stacked 
totalhours_cat_ankle <- totalhours_cat %>% 
  filter(position=="ankle") %>% 
  select(S_ID, position_cathours) %>% 
  rename("ankle_cathours"= "position_cathours")

totalhours_cat_wrist <- totalhours_cat %>% 
  filter(position=="wrist") %>% 
  select(S_ID, position_cathours) %>% 
  rename("wrist_cathours"= "position_cathours")

# only focused on about day x day and night x night not day x night cross for wrist/ankle
totalhours_cat_weartime_matrix <- merge(totalhours_cat_ankle, totalhours_cat_wrist, by="S_ID") %>%
  mutate(combined_total= case_when(wrist_cathours=="wrist high" & ankle_cathours=="ankle high" ~ "high wrist/ high ankle",
                              wrist_cathours=="wrist high" & ankle_cathours=="ankle med" ~ "high wrist/ med ankle",
                              wrist_cathours=="wrist high" & ankle_cathours=="ankle low" ~ "high wrist/ low ankle",
                              wrist_cathours=="wrist med" & ankle_cathours=="ankle high" ~ "med wrist/ high ankle",
                              wrist_cathours=="wrist med" & ankle_cathours=="ankle med" ~ "med wrist/ med ankle",
                              wrist_cathours=="wrist med" & ankle_cathours=="ankle low" ~ "med wrist/ low ankle",
                              wrist_cathours=="wrist low" & ankle_cathours=="ankle high" ~ "low wrist/ high ankle",
                              wrist_cathours=="wrist low" & ankle_cathours=="ankle med" ~ "low wrist/ med ankle",
                              wrist_cathours=="wrist low" & ankle_cathours=="ankle low" ~ "low wrist/ low ankle"))

totalhours_cat_weartime_matrix %>% count(combined_total)
```

# RQ2: TO WHAT EXTENT DO CHILD AND FAMILY CHARACTERISTICS PREDICT CHILD BAND COMPLIANCE?

## a. age and sex
```{r}

demo_weartime <- merge(wk_all_day_hours, totalhours_cat_weartime_matrix)
demo_weartime <- merge(demo_weartime,clean_demo_data_tc)

demo_weartime <- demo_weartime %>% 
  select(-education, # no child education
         -group)     # all children

demo_weartime <- demo_weartime %>% 
  mutate(ankle_cat6day=case_when(ankle_hours < 72 ~ "< 72 hours", 
                                 ankle_hours >= 72 ~ "72 or more hours"), 
         wrist_cat6day=case_when(wrist_hours < 72 ~ "< 72 hours", 
                                 wrist_hours >= 72 ~ "72 or more hours"))

# checks to make sure the categories match 
demo_weartime %>% count(ankle_cat6day)
demo_weartime %>% count(ankle_cathours)
demo_weartime %>% count(wrist_cat6day)
demo_weartime %>% count(wrist_cathours)

# AGE AND DAYTIME HOURS WORN

# plot of age (x) and wrist hours (y)
wristbyage_scatter <-  ggplot(demo_weartime, aes(x=age, y=wrist_hours)) + 
  geom_point(size= 3, color = "orange") + 
  labs(title= "Child Wrist Band Daytime Wear by Age") + 
 ylab("Wrist Daytime Wear") +
  xlab("Age (months)") + 
  stat_cor(method = "spearman", label.x=c(30,30), label.y=c(35,40), digits= 3)

wristbyage_scatter

# plot of age (x) and ankle hours (y)
anklebyage_scatter <-  ggplot(demo_weartime, aes(x=age, y=ankle_hours)) + 
  geom_point(size= 3, color = "cornflowerblue") + 
  labs(title= "Child Ankle Band Daytime Wear by Age") + 
 ylab("Ankle Daytime Wear") +
  xlab("Age (months)") + 
  stat_cor(method = "spearman", label.x=c(30,30), label.y=c(35,40), digits= 3)

anklebyage_scatter



# CHILD SEX AND DAYTIME HOURS WORN
## plots + wilcox rank sum (unpaired) test

# wrist
ggplot(demo_weartime, aes(x=sex, y=wrist_hours, fill= sex)) + 
  geom_boxplot() + 
  labs(title= "Child Wrist Band Daytime Wear by Sex") + 
  theme(legend.position= "none") + 
  scale_fill_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
  geom_jitter(position=position_jitter(0.15)) + 
  stat_compare_means(method= "wilcox.test")

wilcox_test(data= demo_weartime, wrist_hours ~ sex)

# ankle
ggplot(demo_weartime, aes(x=sex, y=ankle_hours, fill= sex)) + 
  geom_boxplot() + 
  labs(title= "Child Ankle Band Daytime Wear by Sex") + 
  theme(legend.position= "none") + 
  scale_fill_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
  geom_jitter(position=position_jitter(0.15)) + 
  stat_compare_means(method= "wilcox.test")

wilcox_test(data= demo_weartime, ankle_hours ~ sex)

# ankle by wrist band scatter (colors = sex)
anklebywrist_scatter_sex <-  ggplot(demo_weartime, aes(x=wrist_hours, y=ankle_hours, color= sex)) + 
  geom_point(size= 3) + 
  labs(title= "Child Band Daytime Wear by Sex") + 
  scale_color_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
 ylab("Ankle Daytime Wear") +
  xlab("Wrist Daytime Wear") + 
  stat_cor(method = "spearman", label.x=c(3,5), label.y=c(30,35))

anklebywrist_scatter_sex



# mean total wear time didn't differ by sex (boxplot/wilcox test) for wrist or ankle bands but for both male and female children, wrist and ankle wear time were correlated (scatter/spearman correlation)



# separate dfs of male and female
demo_weartime_female <- demo_weartime %>% 
  filter(sex=="F")

demo_weartime_male <- demo_weartime %>% 
  filter(sex=="M")


# female only plot
anklebywrist_scatter_f <-  ggplot(demo_weartime_female, aes(x=wrist_hours, y=ankle_hours)) + 
  geom_point(size= 3, color="darkorange") + 
  labs(title= "Child Band Daytime Wear for Female") +
  ylab("Ankle Daytime Wear") +
  xlab("Wrist Daytime Wear") + 
  stat_cor(method = "spearman", label.x=c(3,5), label.y=c(30,35))

anklebywrist_scatter_f

# male only plot
anklebywrist_scatter_m <-  ggplot(demo_weartime_male, aes(x=wrist_hours, y=ankle_hours)) + 
  geom_point(size= 3, color="aquamarine3") + 
  labs(title= "Child Band Daytime Wear for Male") +
  ylab("Ankle Daytime Wear") +
  xlab("Wrist Daytime Wear") + 
  stat_cor(method = "spearman", label.x=c(3,5), label.y=c(30,35), digits= 3)

anklebywrist_scatter_m



# AGE AND DAYTIME HOURS WORN STRATIFIED BY SEX

# ankle hours (y) by age (x) stratified by sex (M= teal, F= orange)
anklebyage_scatter_sex <-  ggplot(demo_weartime, aes(x=age, y=ankle_hours, color= sex)) + 
  geom_point(size= 3) + 
  labs(title= "Child Ankle Band Daytime Wear by Age (x) and Sex (color)") + 
  scale_color_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
 ylab("Ankle Daytime Wear") +
  xlab("Age") + 
  stat_cor(method = "spearman", label.x=c(30,30), label.y=c(35,40))

anklebyage_scatter_sex 

# wrist hours (y) by age (x) stratified by sex (M= teal, F= orange)
wristbyage_scatter_sex <-  ggplot(demo_weartime, aes(x=age, y=wrist_hours, color= sex)) + 
  geom_point(size= 3) + 
  labs(title= "Child Wrist Band Daytime Wear by Age (x) and Sex (color)") + 
  scale_color_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
 ylab("Wrist Daytime Wear") +
  xlab("Age") + 
  stat_cor(method = "spearman", label.x=c(30,30), label.y=c(35,40), digits= 3)

wristbyage_scatter_sex 



demo_weartime_ankle <- demo_weartime %>% 
  select(-wrist_hours, -wrist_cat_hours) %>% 
  mutate(position = "ankle") %>% 
  rename("hours"= "ankle_hours", 
         "cat6day" = "ankle_cat_hours")

demo_weartime_wrist <- demo_weartime %>% 
  select(-ankle_hours, -ankle_cat_hours) %>% 
  mutate(position = "wrist") %>% 
  rename("hours"= "wrist_hours", 
        "cat6day" = "wrist_cat_hours")

demo_weartime_sep <- rbind(demo_weartime_ankle, demo_weartime_wrist)
  


# by position, not paired
weartime_xband_yhours_groupedsex <-  ggplot(demo_weartime_sep, aes(x=position, y=hours, fill= sex)) +
  geom_boxplot(outlier.shape = NA) + 
  labs(title= "Child  Band Daytime Wear by Position and Child Sex", size= 6) + 
  scale_fill_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test",  
                     label.x = 5, label.y = 90, size = 5)

weartime_xband_yhours_groupedsex


# unpaired
ggplot(demo_weartime_sep, aes(x=position, y=hours, fill= position)) + 
  geom_boxplot(outlier.shape = NA) + 
  labs(title= "Child Band Daytime Wear by Position (unpaired)") + 
  scale_fill_manual(values = c("wrist" = "orange", 
                               "ankle" = "cornflowerblue")) +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test",
                      label.y = 90, size = 5)

# paired, not stratified by sex
ggplot(demo_weartime_sep, aes(x=position, y=hours, fill= position)) + 
  geom_boxplot(outlier.shape = NA) + 
  labs(title= "Child Band Daytime Wear by Position (paired)") + 
  scale_fill_manual(values = c("wrist" = "orange", 
                               "ankle" = "cornflowerblue")) +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test", paired= TRUE,
                      label.y = 90, size = 5)


# paired, stratified by sex
weartime_xsex_yhours_groupedband_paired <- ggplot(demo_weartime_sep, aes(x=sex, y=hours, fill= position)) + 
  geom_boxplot(outlier.shape = NA) + 
  labs(title= "Child Band Daytime Wear by Sex (paired)") + 
  scale_fill_manual(values = c("wrist" = "orange", 
                               "ankle" = "cornflowerblue")) +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test", paired= TRUE,
                      label.x = 5, label.y = 90, size = 5)

weartime_xsex_yhours_groupedband_paired

# wilcox rank sum
wilcox.test(hours ~ position, data=demo_weartime_sep)
# wilcox sign rank 
wilcox.test(x = demo_weartime_sep$hours[demo_weartime_sep$position == "wrist"],  
            y = demo_weartime_sep$hours[demo_weartime_sep$position == "ankle"],  
            paired = TRUE)

# ** results presented under RQ1 results section manuscript **


# mean daytime wear hours for ankle and wrist
demo_weartime_sep %>% group_by(position) %>% summarise(mean(hours))

# mean daytime wear hours for ankle and wrist stratified by sex 
demo_weartime_sep %>% group_by(position, sex) %>% summarise(mean(hours))

```

## b. band wear time categories (total hours) and age
```{r}
# visual check for normality 
hist(demo_weartime$age)

# quant check for normality
shapiro.test(demo_weartime$age) # p= 0.06219, age is normally distributed, ok to use anova

demo_weartime$ankle_cathours <- factor(demo_weartime$ankle_cathours, 
                                    levels=c("ankle low", "ankle med", "ankle high"), ordered=TRUE)  

demo_weartime$wrist_cathours <- factor(demo_weartime$wrist_cathours, 
                                    levels=c("wrist low", "wrist med", "wrist high"), ordered=TRUE)

# ANOVA TO COMPARE THREE GROUPS (HIGH MED LOW)

# ankle band
ankle_anova <- aov(age ~ ankle_cathours,
            data= demo_weartime)
summary(ankle_anova)
    # no post hoc test needed

# wrist band
wrist_anova <- aov(age ~ wrist_cathours,
            data= demo_weartime) 
summary(wrist_anova)

# post hoc test for multiple comparisons
tukey_hsd(wrist_anova, conf.level=0.95)

    # age of high wearers is sig dif (adj p= 0.0275)

demo_weartime %>% group_by(wrist_cathours) %>% summarise(mean(age))


# PLOTS

# ankle band box
ggplot(demo_weartime, aes(x=ankle_cathours, y=age, fill=ankle_cathours)) + geom_boxplot() + labs(title= "Boxplot of Ankle Cat by Age") + theme(legend.position= "none")

# wrist band box
ggplot(demo_weartime, aes(x=wrist_cathours, y=age, fill=wrist_cathours)) + geom_boxplot() + labs(title= "Boxplot of Wrist Cat by Age") + theme(legend.position= "none")
```

## c. long sleeves/ pants
```{r}

# selecting variables with sleeve/ pant data
sleeves_pants_day <- bands %>% 
  select(S_ID, 
         contains("sleeve"), 
         contains("pant"),
         contains("daytime"), 
         contains("week")) %>%  
  rename("child_wrist_day1" = "bwb_cwhrsdaytime1_T30", 
         "child_ankle_day1" = "bwb_cahrsdaytime1_T30", 
         "parent_wrist_day1"= "bwb_pwhrsdaytime1_T30",
         "child_wrist_day2" = "bwb_cwhrsdaytime2_T30",
         "child_ankle_day2" = "bwb_cahrsdaytime2_T30",
         "parent_wrist_day2" = "bwb_pwhrsdaytime2_T30",
         "child_wrist_day3" = "bwb_cwhrsdaytime3_T30", 
         "child_ankle_day3" = "bwb_cahrsdaytime3_T30", 
         "parent_wrist_day3" = "bwb_pwhrsdaytime3_T30",
         "child_wrist_day4" = "bwb_cwhrsdaytime4_T30", 
         "child_ankle_day4" = "bwb_cahrsdaytime4_T30", 
         "parent_wrist_day4" = "bwb_pwhrsdaytime4_T30",
         "child_wrist_day5" = "bwb_cwhrsdaytime5_T30", 
         "child_ankle_day5" = "bwb_cahrsdaytime5_T30", 
         "parent_wrist_day5" = "bwb_pwhrsdaytime5_T30",
         "child_wrist_day6" = "bwb_cwhrsdaytime6_T30", 
         "child_ankle_day6" = "bwb_cahrsdaytime6_T30", 
         "parent_wrist_day6" = "bwb_pwhrsdaytime6_T30",
         "child_wrist_day7" = "bwb_cwhrsdaytime7_T30", 
         "child_ankle_day7" ="bwb_cahrsdaytime7_T30", 
         "parent_wrist_day7" = "bwb_pwhrsdaytime7_T30",
         
         "child_wrist_wktotal" = "bwb_cwhrsweek_T30",
         "child_wrist_daytotal" = "bwb_cwhrsweekday_T30",
         "child_wrist_nighttotal" = "bwb_cwhrsweeknite_T30",
         "child_ankle_wktotal" = "bwb_cahrsweek_T30",
         "child_ankle_daytotal" = "bwb_cahrsweekday_T30",
         "child_ankle_nighttotal" = "bwb_cahrsweeknite_T30",
         
         "day1_sleeves" = "bwb_day1_cwsleeves_T30",
         "day2_sleeves" = "bwb_day2_cwsleeves_T30", 
         "day3_sleeves" = "bwb_day3_cwsleeves_T30",
         "day4_sleeves" = "bwb_day4_cwsleeves_T30", 
         "day5_sleeves" = "bwb_day5_cwsleeves_T30", 
         "day6_sleeves" = "bwb_day6_cwsleeves_T30", 
         "day7_sleeves" = "bwb_day7_cwsleeves_T30", 
         
         "day1_pants" = "bwb_day1_capants_T30",
         "day2_pants" = "bwb_day2_capants_T30",
         "day3_pants" = "bwb_day3_capants_T30", 
         "day4_pants" = "bwb_day4_capants_T30", 
         "day5_pants" = "bwb_day5_capants_T30", 
         "day6_pants" = "bwb_day6_capants_T30", 
         "day7_pants" = "bwb_day7_capants_T30")

sleeves_pants_day_long <- sleeves_pants_day %>% 
  select(S_ID, contains("sleeve"), contains("pants")) %>% 
  pivot_longer(!S_ID, names_to = "longname", values_to= "present")

sleeves_wrist <- sleeves_pants_day %>% 
  select(S_ID, (contains("wrist") & contains("child")), contains("sleeve")) %>% 
  mutate(n.na=rowSums(is.na(select(.,c(day1_sleeves:day7_sleeves)))), 
         n.yes=rowSums(select(.,c(day1_sleeves:day7_sleeves)) == 1, na.rm=TRUE),
         n.no=rowSums((select(.,c(day1_sleeves:day7_sleeves))) == 0, na.rm=TRUE)) %>% 
  rowwise() %>% 
  mutate(n.total=sum(c_across(c(n.na:n.no))),
         sleeves_per_yes= n.yes / 7) %>% #check
  ungroup()


# without removing rows with all missing
sleeves_percent_scatter_all <- ggplot(sleeves_wrist, aes(x=sleeves_per_yes, y=child_wrist_daytotal)) + 
  geom_point(color="orange") + 
  stat_cor(method = "spearman", label.y=c(30,35), digits= 3) +
  labs(title= "Daytime Wristband Hours by Percent of the week wearing sleeves") + 
  ylab("Total Daytime hours") + 
  xlab("Percent of the week wearing sleeves")
  
sleeves_percent_scatter_all 

#removing the participants that had n.na=7 (all missing)
sleeves_wrist <- sleeves_wrist %>% 
  filter(!n.na==7)

sleeves_percent_scatter_noallna <- ggplot(sleeves_wrist, aes(x=sleeves_per_yes, y=child_wrist_daytotal)) + 
  geom_point(color="orange") + 
  stat_cor(method = "spearman", label.y=c(30,35), digits= 3) +
  labs(title= "Daytime Wristband Hours by Percent of the week wearing sleeves") + 
  ylab("Total Daytime hours") + 
  xlab("Percent of the week wearing sleeves")
  
sleeves_percent_scatter_noallna 

pants_ankle <- sleeves_pants_day %>% 
  select(S_ID, contains("ankle"), contains("pant")) %>% 
  mutate(n.na=rowSums(is.na(select(.,c(day1_pants:day7_pants)))), 
         n.yes=rowSums(select(.,c(day1_pants:day7_pants)) == 1, na.rm=TRUE),
         n.no=rowSums((select(.,c(day1_pants:day7_pants))) == 0, na.rm=TRUE)) %>% 
  rowwise() %>% 
  mutate(n.total=sum(c_across(c(n.na:n.no))),
         pants_per_yes= n.yes / 7) %>% #check
  ungroup()

pants_percent_scatter_all <- ggplot(pants_ankle, aes(x=pants_per_yes, y=child_ankle_daytotal)) + 
  geom_point(color="cornflowerblue") + 
  stat_cor(method = "spearman", label.y=c(30,35)) +
  labs(title= "Daytime Ankle Band Hours by Percent of the week wearing sleeves") + 
  ylab("Total Daytime hours") + 
  xlab("Percent of the week wearing pants")

pants_percent_scatter_all

# removing the participants that had n.na=7 (all missing)
pants_ankle <- pants_ankle %>% 
  filter(!n.na==7)

pants_percent_scatter_noallna <- ggplot(pants_ankle, aes(x=pants_per_yes, y=child_ankle_daytotal)) + 
  geom_point(color="cornflowerblue") + 
  stat_cor(method = "spearman", label.y=c(30,35), digits= 3) +
  labs(title= "Daytime Ankle Band Hours by Percent of the week wearing sleeves") + 
  ylab("Total Daytime hours") + 
  xlab("Percent of the week wearing pants")

pants_percent_scatter_noallna

```

## d. childcare
### ccare1_T30: Childcare Q1. Typical week, how many places does child receive care outside of self and live-in partner?
```{r}

# look at counts for how many places the child receives care outside the home in a typical week (range = 0-5)
bands %>% count(ccare1_T30)

# categorize to collapse small categories 
childcare <- bands %>% 
  select(S_ID, ccare1_T30) %>% 
  mutate(childcare_cat = case_when(ccare1_T30== 0 ~ "0",             
         # 3 categories
                                   ccare1_T30== 1 ~ "1", 
                                   ccare1_T30== 2 ~ "2 or more", 
                                   ccare1_T30== 3 ~ "2 or more", 
                                   ccare1_T30== 4 ~ "2 or more",
                                   ccare1_T30== 5 ~ "2 or more"),
         childcare_bi = case_when(ccare1_T30== 0 ~ "No chilcare",     
         # binary, childcare/no
                                   ccare1_T30== 1 ~ "Childcare", 
                                   ccare1_T30== 2 ~ "Childcare", 
                                   ccare1_T30== 3 ~ "Childcare", 
                                   ccare1_T30== 4 ~ "Childcare",
                                   ccare1_T30== 5 ~ "Childcare"), 
         childcare_cat_num = case_when(ccare1_T30== 0 ~ 0,            
         # 3 categories, as numeric for anova
                                   ccare1_T30== 1 ~ 1, 
                                   ccare1_T30== 2 ~ 2, 
                                   ccare1_T30== 3 ~ 2, 
                                   ccare1_T30== 4 ~ 2,
                                   ccare1_T30== 5 ~ 2))
# checks
childcare %>% count(childcare_cat)
childcare %>% count(childcare_cat_num)
childcare %>% count(childcare_bi)
# counts match the noncollapsed vars 


# merge with wear time
table3_childcare <- merge(childcare, demo_weartime)
# table3_childcare %>% group_by(wrist_cat6day) %>% count(childcare_cat)
# table3_childcare %>% group_by(ankle_cat6day) %>% count(childcare_cat)

hist(table3_childcare$wrist_hours) # non normal 




# WRIST

# kruskal wallis test for multiple comparisons (non normal data)
kruskal.test(wrist_hours ~ childcare_cat, data= table3_childcare)

# plot

# 3 group comparison
childcare_wrist_box <- ggplot(table3_childcare, aes(x=childcare_cat, y=wrist_hours, fill=childcare_cat)) + 
  geom_boxplot(alpha=1, fill= "orange", outlier.shape=NA) +
  stat_kruskal_test(label= "Kruskal-Wallis, p = {p.format}", label.y= 86) +
  geom_jitter(size=0.5, alpha= 0.9, position = position_jitterdodge(0.07)) +
  labs(title= "Child Wrist Band Total Wear Times by Number of Child Care facilities")

childcare_wrist_box

# 2 group comparison
childcarebi_wrist_box <- ggplot(table3_childcare, aes(x=childcare_bi, y=wrist_hours, fill=childcare_bi)) + 
  geom_boxplot(alpha=1, fill= "orange", outlier.shape=NA) + 
  geom_jitter(size=0.5, alpha= 0.9, position = position_jitterdodge(0.07)) + 
  labs(title= "Child Wrist Band Total Wear Times by Number of Child Care facilities") +
  stat_compare_means(method= "wilcox.test", label.y = 90)

childcarebi_wrist_box 




# ANKLE

hist(table3_childcare$ankle_hours) # non normal 

# kruskal wallis test for multiple comparisons (non normal data)
kruskal.test(ankle_hours ~ childcare_cat, data= table3_childcare)

# plot

# 3 group comparison
childcare_ankle_box <- ggplot(table3_childcare, aes(x=childcare_cat, y=ankle_hours, fill=childcare_cat)) + 
  geom_boxplot(alpha=1, fill= "cornflowerblue", outlier.shape=NA) + 
  stat_kruskal_test(label= "Kruskal-Wallis, p = {p.format}", label.y= 86) +
  geom_jitter(size=0.5, alpha= 0.9, position = position_jitterdodge(0.07)) + 
  labs(title= "Child Ankle Band Total Wear Times by Number of Child Care facilities")

childcare_ankle_box

# 2 group comparison
childcarebi_ankle_box <- ggplot(table3_childcare, aes(x=childcare_bi, y=ankle_hours, fill=childcare_bi)) + 
  geom_boxplot(alpha=1, fill= "cornflowerblue", outlier.shape=NA) + 
  geom_jitter(size=0.5, alpha= 0.9, position = position_jitterdodge(0.07)) + 
  labs(title= "Child Ankle Band Total Wear Times by Number of Child Care facilities") + 
  stat_compare_means(method= "wilcox.test", label.y = 90)

childcarebi_ankle_box



```

## e. disabilities
### cdc Does child have any suspected or identified disabilities?
```{r}

# select variable and merge with wear time 
## 
cdc <- bands %>% 
  select(S_ID, cdc_disany_T30) 

cdc <- merge(cdc, wk_child_day_hours)

cdc %>% count(cdc_disany_T30)
  # 0 == none       N=83
  # 1 == yes        N=19
  # NA == missing   N=13

cdc <- cdc %>% 
  filter(!is.na(cdc_disany_T30)) %>% 
  mutate(disany_binary=case_when(cdc_disany_T30 == 0 ~ "none", 
                                 cdc_disany_T30 == 1 ~ "one"))

# WRIST 
cdc_wrist_wilcox <- wilcox_test(data=cdc, wrist_hours ~ disany_binary)
formatted_stat <- format.pval(cdc_wrist_wilcox$statistic, digits = 3)
print(formatted_stat)

# ANKLE
cdc_ankle_wilcox <- wilcox_test(data=cdc, ankle_hours ~ disany_binary)
formatted_stat <- format.pval(cdc_ankle_wilcox$statistic, digits = 3)
print(formatted_stat)
```

## f. spearman correlations between attributes and child wrist and ankle wear time
```{r}

# WRIST
attributes <- as.vector(colnames(wk_child_day_wrist[5:15]))
is.vector(attributes)

# empty df to store results
results_df <- data.frame(
  variable = character(),
  correlation = numeric(),
  p_value = numeric(),
  conf_lower = numeric(),
  conf_upper = numeric(),
  stringsAsFactors = FALSE
)

# spearman correlation for each variable in the list
for (var in attributes) {
  test_result <- cor.test(wk_child_day_wrist[[var]], wk_child_day_wrist$hours, method= "spearman")
  
  p_value <- ifelse(is.na(test_result$p.value), NA, test_result$p.value)
  
  # append the results to the dataframe
  results_df <- rbind(results_df, data.frame(
    variable = var,
    correlation = test_result$estimate,
    p_value = p_value,
    stringsAsFactors = FALSE
  ))
}

# rename 
corr_child_atttributesdaywear_wrist <- results_df

# check for psi_mn
cor.test(wk_child_day_wrist$PSI_Mn_T30, wk_child_day_wrist$hours, method= "spearman")




# ANKLE
attributes <- as.vector(colnames(wk_child_day_ankle[5:15]))
is.vector(attributes)

# empty df to store results
results_df <- data.frame(
  variable = character(),
  correlation = numeric(),
  p_value = numeric(),
  conf_lower = numeric(),
  conf_upper = numeric(),
  stringsAsFactors = FALSE
)

# spearman correlation for each variable in the list
for (var in attributes) {
  test_result <- cor.test(wk_child_day_ankle[[var]], wk_child_day_ankle$hours, method= "spearman")
  
  p_value <- ifelse(is.na(test_result$p.value), NA, test_result$p.value)
  
  # append the results to the dataframe
  results_df <- rbind(results_df, data.frame(
    variable = var,
    correlation = test_result$estimate,
    p_value = p_value,
    stringsAsFactors = FALSE
  ))
}

# rename 
corr_child_atttributesdaywear_ankle <- results_df

# check
cor.test(wk_child_day_ankle$PSI_Mn_T30, wk_child_day_ankle$hours, method= "spearman")

```

## g. wilcox rank sum tests to see if wear time groups have dif attribute scores (low/mod; less than 72 hours vs. high ≥72 hours)
### table 3
```{r}

# modified from TAME2 Ch4 "Unpaired example" https://uncsrp.github.io/TAME2/two-group-comparisons-and-visualizations.html#unpaired-test-example

# WRIST 
endpoints <- colnames(wk_child_day_wrist %>% select(PSI_Mn_T30:PAS_AVG_T30)) 

# create dataframe to store results
cat_hours_wrist_wilcoxres <- data.frame()

# run for loop
for (i in 1:length(endpoints)) {
  
  # assign a name to the endpoint variable.
  endpoint <- endpoints[i]
  
  # run wilcox test and store in results dataframe.
  res_df <- wk_child_day_wrist %>%
    wilcox_test(as.formula(paste0(endpoint, "~ cat_hours", sep = "")))
  
  # bind results from this test with other tests in this loop
  cat_hours_wrist_wilcoxres <- rbind(cat_hours_wrist_wilcoxres, res_df)
  
}

cat_hours_wrist_wilcoxres

#ANKLE
endpoints <- colnames(wk_child_day_ankle %>% select(PSI_Mn_T30:PAS_AVG_T30)) 

# create dataframe to store results
cat_hours_ankle_wilcoxres <- data.frame()

# run for loop
for (i in 1:length(endpoints)) {
  
  # assign a name to the endpoint variable.
  endpoint <- endpoints[i]
  
  # run wilcox test and store in results dataframe.
  res_df <- wk_child_day_ankle %>%
    wilcox_test(as.formula(paste0(endpoint, "~ cat_hours", sep = ""))) 
  
  # bind results from this test with other tests in this loop
  cat_hours_ankle_wilcoxres <- rbind(cat_hours_ankle_wilcoxres, res_df)
  
}

# manual look at the means 
wk_child_day_ankle %>% 
  filter(cat_hours=="< 72 hours") %>% 
  summarise(mean(ECBQ_EFFCO_T30, na.rm=TRUE))

wk_child_day_ankle %>% 
  filter(cat_hours=="72 or more hours") %>% 
  summarise(mean(ECBQ_EFFCO_T30, na.rm=TRUE))

```

## h. additional child characteristics (sex, childcare) and parent attributes (income to needs, education, age, stress)
### table 3
```{r}

# df of demo and additional attributes  
additional_attributes <- bands %>% 
  select(S_ID, 
         TC_AgeMonths_Home_T30, 
         MCDI3_All_T30, #all 
         MCDI3_Use_T30, #language only
         PAS_AVG_T30,   #total anxiety
         BIQ_Soc_T30,   
         BIQ_Nov_T30, 
         BIQ_Tot_T30,   #total 
         ECBQ_dis_T30, 
         ECBQ_NEGAF_T30, 
         ECBQ_SURGE_T30, 
         ECBQ_EFFCO_T30,
         income2needs_T30, 
         PC_matage_home_T30,
         pc_ed,
         PSI_Mn_T30,   #mean across all items
         PSI_DC_T30)

additional_attributes <- merge(demo_weartime, additional_attributes, by="S_ID")

additional_attributes <- additional_attributes %>% 
  select(-latino_hispanic, 
         -race) %>% 
  relocate(adult_hours, .after=PSI_DC_T30)

# WILCOX RANK SUM for WRIST
endpoints <- colnames(additional_attributes %>% select(TC_AgeMonths_Home_T30:adult_hours)) 

# create dataframe to store results
addatt_wrist_wilcoxres <- data.frame()

# run for loop
for (i in 1:length(endpoints)) {
  
  # assign a name to the endpoint variable.
  endpoint <- endpoints[i]
  
  # calculate means for the two groups
  means <- additional_attributes %>%
    group_by(wrist_cat6day) %>%
    summarise(mean_value = mean(get(endpoint), na.rm = TRUE)) %>%
    pivot_wider(names_from = wrist_cat6day, values_from = mean_value, names_prefix = "mean_")
  
  # run Wilcoxon test and store in results dataframe.
  res_df <- additional_attributes %>%
    wilcox_test(as.formula(paste0(endpoint, "~ wrist_cat6day", sep = ""))) %>%
    left_join(means, by = character()) %>% 
    mutate(across(where(is.numeric), ~ round(.x, 3)))
  
  # bind results from this test with other tests in this loop
  addatt_wrist_wilcoxres <- rbind(addatt_wrist_wilcoxres, res_df)
  
}

addatt_wrist_wilcoxres

is.numeric(addatt_wrist_wilcoxres$statistic)

# WILCOX RANK SUM for ANKLE

# create dataframe to store results
addatt_ankle_wilcoxres <- data.frame()

# run for loop
for (i in 1:length(endpoints)) {
  
  # assign a name to the endpoint variable.
  endpoint <- endpoints[i]
  
  # calculate means for the two groups
  means <- additional_attributes %>%
    group_by(ankle_cat6day) %>%
    summarise(mean_value = mean(get(endpoint), na.rm = TRUE)) %>%
    pivot_wider(names_from = ankle_cat6day, values_from = mean_value, names_prefix = "Mean_")
  
  # run Wilcoxon test and store in results dataframe.
  res_df <- additional_attributes %>%
    wilcox_test(as.formula(paste0(endpoint, "~ ankle_cat6day", sep = ""))) %>%
    mutate(Variable = endpoint) %>%
    left_join(means, by = character())
  
  # bind results from this test with other tests in this loop
  addatt_ankle_wilcoxres <- rbind(addatt_ankle_wilcoxres, res_df)
  
}

addatt_ankle_wilcoxres

addatt_ankle_wilcoxres$statistic <- round(addatt_ankle_wilcoxres$statistic, digits = 3)

# WRIST
# check with the counts
additional_attributes %>% group_by(wrist_cat6day) %>% count(wrist_cat6day)
# mean age of the two groups
additional_attributes %>% group_by(wrist_cat6day) %>% summarise(mean= mean(age))
# sex distribution within the two groups
additional_attributes %>% group_by(wrist_cat6day) %>% count(sex)

# ANKLE
# check 
additional_attributes %>% group_by(ankle_cat6day) %>% count(ankle_cat6day)
# mean age of the two groups
additional_attributes %>% group_by(ankle_cat6day) %>% summarise(mean= mean(age))
# sex distribution within the two groups
additional_attributes %>% group_by(ankle_cat6day) %>% count(sex)



## SEX IS CATEGORICAL VAR ##

# assessing independence in the weartime groups by sex
chisq_ankle_sex <- chisq.test(additional_attributes$sex, additional_attributes$ankle_cat_hours)
format(round(chisq_ankle_sex$p.value, 3), nsmall = 3)

chisq_wrist_sex <-chisq.test(additional_attributes$sex, additional_attributes$wrist_cat_hours)
format(round(chisq_wrist_sex$p.value, 3), nsmall = 3)


# creating new dataframe to do wilcox test comparing mean use of long sleeves/pants over the 7 day week
sleeves_addatt <- sleeves_wrist %>% 
  select(S_ID, sleeves_per_yes)

sleeves_addatt <- merge(sleeves_addatt, additional_attributes, by="S_ID")

sleeves_wilcox <- wilcox.test(data=sleeves_addatt, sleeves_per_yes ~ wrist_cat6day)
format(round(sleeves_wilcox$p.value, 3), nsmall = 3)


pants_addatt <- pants_ankle %>% 
  select(S_ID, pants_per_yes)

pants_addatt <- merge(pants_addatt, additional_attributes, by="S_ID")

pants_wilcox <- wilcox.test(data=pants_addatt, pants_per_yes ~ ankle_cat6day)
format(round(pants_wilcox$p.value, 3), nsmall = 3)

### CORRELATIONS ###

# parental stress and parent wear time (only wrist)
# mean of PSI
cor.test(wk_adult_day$PSI_Mn_T30, wk_adult_day$hours, method= "spearman")
# difficult child
cor.test(wk_adult_day$PSI_DC_T30, wk_adult_day$hours, method= "spearman")

## child wrist and ankle wear time correlations with PSI are in respective dfs (corr_child_atttributesdaywear_ankle/ corr_child_atttributesdaywear_wrist)



### INCOME TO NEEDS
# wrist 
cor.test(additional_attributes$income2needs_T30, additional_attributes$wrist_hours, method="spearman")

# ankle
cor.test(additional_attributes$income2needs_T30, additional_attributes$ankle_hours, method="spearman")



### mat edu
# wrist 
cor.test(additional_attributes$pc_ed, additional_attributes$wrist_hours, method="spearman")

# ankle 
cor.test(additional_attributes$pc_ed, additional_attributes$ankle_hours, method="spearman")



### MAT AGE
# wrist 
cor.test(additional_attributes$PC_matage_home_T30, additional_attributes$wrist_hours, method="spearman")

# ankle
cor.test(additional_attributes$PC_matage_home_T30, additional_attributes$ankle_hours, method="spearman")

```

## i. linear regression
### 1. unadjusted (only attribute) 
### 2. attribute and adjusted for age and sex
```{r}

regressiondf <- bands %>%  # attributes we found to be significantly different at p<0.05 between the low/mod and high wear time groups
  select(S_ID, 
         MCDI3_All_T30,    # wrist
         PAS_AVG_T30,      # wrist
         BIQ_Tot_T30,      # ankle
         BIQ_Nov_T30,      # ankle
         BIQ_Soc_T30,      # ankle
         ECBQ_EFFCO_T30)   # ankle

demo_weartime_reg <- demo_weartime %>% 
  select(S_ID, sex, age, wrist_hours, ankle_hours)

regressiondf <- merge(regressiondf, demo_weartime_reg, by="S_ID")

## UNADJUSTED
# WRIST

## MCDI
regression_wrist_mcdi <- lm(wrist_hours ~ MCDI3_All_T30,
                       data=regressiondf)
regression_wrist_mcdi_crudesumm <- summary(regression_wrist_mcdi)
regression_wrist_mcdi_conf <- confint(regression_wrist_mcdi, level= 0.95)

regression_wrist_mcdi_crudesumm
regression_wrist_mcdi_conf

## PAS
regression_wrist_pas <- lm(wrist_hours ~ PAS_AVG_T30, 
                       data=regressiondf)
regression_wrist_pas_crudesumm <- summary(regression_wrist_pas)
regression_wrist_pas_conf <- confint(regression_wrist_pas, level= 0.95)

regression_wrist_pas_crudesumm
regression_wrist_pas_conf



# ANKLE 

## ECBQ Effortful control
regression_ankle_ecbq <- lm(ankle_hours ~ ECBQ_EFFCO_T30, 
                       data=regressiondf)
regression_ankle_ecbq_crudesumm <- summary(regression_ankle_ecbq)
regression_ankle_ecbq_conf <- confint(regression_ankle_ecbq, level= 0.95)

regression_ankle_ecbq_crudesumm
regression_ankle_ecbq_conf

## BIQ social 
regression_ankle_biqsoc <- lm(ankle_hours ~ BIQ_Soc_T30, 
                       data=regressiondf)
regression_ankle_biqsoc_crudesumm <- summary(regression_ankle_biqsoc)
regression_ankle_biqsoc_conf <- confint(regression_ankle_biqsoc, level= 0.95)

regression_ankle_biqsoc_crudesumm
regression_ankle_biqsoc_conf

## BIQ novelty
regression_ankle_biqnov <- lm(ankle_hours ~ BIQ_Nov_T30, 
                       data=regressiondf)
regression_ankle_biqnov_crudesumm <- summary(regression_ankle_biqnov)
regression_ankle_biqnov_conf <- confint(regression_ankle_biqnov, level= 0.95)

regression_ankle_biqnov_crudesumm
regression_ankle_biqnov_conf

## BIQ total
regression_ankle_biqtot <- lm(ankle_hours ~ BIQ_Tot_T30, 
                       data=regressiondf)
regression_ankle_biqtot_cudesumm <- summary(regression_ankle_biqtot)
regression_ankle_biqtot_conf <- confint(regression_ankle_biqtot, level= 0.95)

regression_ankle_biqtot_cudesumm
regression_ankle_biqtot_conf






## ADJUSTED
# WRIST

## MCDI
regression_wrist_mcdi <- lm(wrist_hours ~ MCDI3_All_T30 + age + sex,
                       data=regressiondf)
regression_wrist_mcdi_adjsumm <- summary(regression_wrist_mcdi)
regression_wrist_mcdi_adjconf <- confint(regression_wrist_mcdi, level= 0.95)

regression_wrist_mcdi_adjsumm
regression_wrist_mcdi_adjconf

## PAS
regression_wrist_pas <- lm(wrist_hours ~ PAS_AVG_T30 + age + sex, 
                       data=regressiondf)
regression_wrist_pas_adjsumm <- summary(regression_wrist_pas)
regression_wrist_pas_adjconf <- confint(regression_wrist_pas, level= 0.95)

regression_wrist_pas_adjsumm
regression_wrist_pas_adjconf



# ANKLE 

## ECBQ Effortful control
regression_ankle_ecbq <- lm(ankle_hours ~ ECBQ_EFFCO_T30 + age + sex, 
                       data=regressiondf)
regression_ankle_ecbq_adjsumm <- summary(regression_ankle_ecbq)
regression_ankle_ecbq_adjconf <- confint(regression_ankle_ecbq, level= 0.95)

regression_ankle_ecbq_adjsumm
regression_ankle_ecbq_adjconf

## BIQ social 
regression_ankle_biqsoc <- lm(ankle_hours ~ BIQ_Soc_T30 + age + sex, 
                       data=regressiondf)
regression_ankle_biqsoc_adjsumm <- summary(regression_ankle_biqsoc)
regression_ankle_biqsoc_adjconf <- confint(regression_ankle_biqsoc, level= 0.95)

regression_ankle_biqsoc_adjsumm
regression_ankle_biqsoc_adjconf

## BIQ novelty
regression_ankle_biqnov <- lm(ankle_hours ~ BIQ_Nov_T30 + age + sex, 
                       data=regressiondf)
regression_ankle_biqnov_adjsumm <- summary(regression_ankle_biqnov)
regression_ankle_biqnov_adjconf <- confint(regression_ankle_biqnov, level= 0.95)

regression_ankle_biqnov_adjsumm
regression_ankle_biqnov_adjconf

## BIQ total
regression_ankle_biqtot <- lm(ankle_hours ~ BIQ_Tot_T30 + age + sex, 
                       data=regressiondf)
regression_ankle_biqtot_adjsumm <- summary(regression_ankle_biqtot)
regression_ankle_biqtot_adjconf <- confint(regression_ankle_biqtot, level= 0.95)

regression_ankle_biqtot_adjsumm
regression_ankle_biqtot_adjconf

```

# RQ3: WHAT RECOMMENDATIONS CAN WE PROVIDE TO IMPROVE COMPLIANCE AND PARTICIPANT EXPERIENCES FOR FUTURE STUDIES?

## a. issues
```{r}

# issues with the bands from the qualitative, write-in question for the band diary
issues_nobwb <- issues %>% 
  select(-contains("bwb"))

demo_issues_weartime <- merge(demo_weartime, issues_nobwb, by="S_ID", all.x= TRUE) 

demo_issues_weartime_long <- demo_issues_weartime %>% 
  filter(!is.na(Size)) %>% # those with missing for size, missing for all
  select(-ankle_cat6day, -wrist_cat6day, 
         -adult_hours, -latino_hispanic, -race, -ankle_cat_hours, -wrist_cat_hours) %>% #relevant variables
  pivot_longer(!c(S_ID, ankle_hours, wrist_hours, combined_total, ankle_cathours, wrist_cathours, age, sex), names_to = "issue", values_to = "present") 

sum_issues <- demo_issues_weartime_long %>% 
  group_by(issue) %>% 
  summarise(sum=sum(present))

issues_long <- issues_nobwb %>% 
  pivot_longer(!S_ID, names_to = "type", values_to = "present") %>% 
  filter(present==1) #to include only those that had an issue (no 0s)


sum_issues_bar <- issues_long %>% 
  ggplot(aes(x=type)) +
  geom_bar(stat= "count", fill= "darkgreen") 

sum_issues_bar


demo_issues_weartime <- demo_issues_weartime %>% 
  filter(!is.na(Size)) %>% 
  select(-ankle_cat_hours, -wrist_cat_hours, -wrist_cathours, -ankle_cathours)

# are the same people leaving comments?
demo_issues_weartime$composite_comments <- rowSums(demo_issues_weartime[, c(12:18, 20:21)], na.rm=TRUE) #column 19 is no notes
demo_issues_weartime$composite_nopos <- rowSums(demo_issues_weartime[, c(12:15,17, 18, 20, 21)], na.rm=TRUE)
                                                                            #removes column 16 (positive) and 19 (no notes)

demo_issues_weartime %>% count(composite_comments) #including the positive comments

demo_issues_weartime %>% count(composite_nopos) #not including the positive comments, only issues

# sample size for 3 and 4 comments is a bit small, going to look at wear time based on 0, 1, and 2 or more comments
demo_issues_weartime <- demo_issues_weartime %>% 
  mutate(comment_cat = case_when(composite_nopos == 0 ~ "0", 
                                composite_nopos == 1 ~ "1",
                                composite_nopos == 2 ~ "2 or more", 
                                  composite_nopos == 3 ~ "2 or more",
                                  composite_nopos == 4  ~ "2 or more"
                                ))
demo_issues_weartime %>% 
  count(comment_cat)


# issues to wrist hours
issues_wrist_box <- ggplot(demo_issues_weartime, aes(x=comment_cat, y=wrist_hours, fill=comment_cat)) +
  geom_boxplot( outlier.shape = NA, fill= "orange") +
  labs(title= "Daytime Wristband Wear by Number of Issues") +
  ylab("Day Time Wrist Hours") +
  xlab("Number of Issues") +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) + 
  stat_kruskal_test(label= "Kruskal-Wallis, p = {p.format}", label.y= 86) 
 
issues_wrist_box

# kruskal wallis test for multiple comparisons
kruskal.test(wrist_hours ~ comment_cat, data= demo_issues_weartime)

# post hoc test for multiple comparisons (dunn's test)
dunnTest(wrist_hours ~ comment_cat, data=demo_issues_weartime, 
         method="bonferroni")


issues_ankle_box <- ggplot(demo_issues_weartime, aes(x=comment_cat, y=ankle_hours, fill=comment_cat)) +
  geom_boxplot( outlier.shape = NA, fill= "cornflowerblue") +
  labs(title= "Daytime Ankleband Wear by Number of Issues") +
  ylab("Day Time Ankle Hours") +
  xlab("Number of Issues") +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) + 
  stat_kruskal_test(label= "Kruskal-Wallis, p = {p.format}", label.y= 86) 
 
issues_ankle_box

kruskal.test(ankle_hours ~ comment_cat, data= demo_issues_weartime)

# post hoc test for multiple comparisons (dunn's test)
dunnTest(ankle_hours ~ comment_cat, data=demo_issues_weartime, 
         method="bonferroni")






ankle_issues_weartime <- demo_issues_weartime %>% 
  select(-wrist_hours) %>% #keep everything but the wrist wear time
  rename("hours"= "ankle_hours") %>% 
  mutate(position= "ankle")

wrist_issues_weartime <- demo_issues_weartime %>% 
  select(-ankle_hours) %>%
  rename("hours"= "wrist_hours") %>% 
  mutate(position= "wrist")

position_issues_weartime <- rbind(ankle_issues_weartime, wrist_issues_weartime)

## paired
issues_byposition <- ggplot(position_issues_weartime, aes(x=comment_cat, y=hours, fill=position)) +
  geom_boxplot(alpha= 1, outlier.shape = NA) +
  scale_fill_manual(values = c("wrist" = "orange", 
                               "ankle" = "cornflowerblue")) +
  labs(title= "Daytime Band Wear by Number of Issues") +
  ylab("Day Time Hours") +
  xlab("Number of Comments") +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test", paired = TRUE,
                      label.y = 90, size = 5)

issues_byposition


issues_wrist_bysex <- ggplot(demo_issues_weartime, aes(x=comment_cat, y=wrist_hours, fill=sex)) +
  geom_boxplot(alpha= 1, outlier.shape = NA) +
  scale_fill_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
  labs(title= "Daytime Wristband Wear by Number of Issues (by sex)") +
  ylab("Day Time Wrist Hours") +
  xlab("Number of Comments") +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test",
                      label.y = 90, size = 5)
  
issues_wrist_bysex


issues_ankle_bysex <- ggplot(demo_issues_weartime, aes(x=comment_cat, y=ankle_hours, fill=sex)) +
  geom_boxplot(alpha= 1, outlier.shape = NA) +
  scale_fill_manual(values = c("F" = "darkorange", 
                               "M" = "aquamarine3")) +
  labs(title= "Daytime Ankle Band Wear by Number of Issues (by sex)") +
  ylab("Day Time Ankle Hours") +
  xlab("Number of Comments") +
  geom_jitter(size=0.3, alpha= 0.9, position = position_jitterdodge(0.07)) +
  stat_compare_means(method = "wilcox.test",
                      label.y = 90, size = 5)
  
issues_ankle_bysex

```


Random forest code by allison spring 
```{r}

set.seed(12)

# random forest
randomforest_demo_data <- data %>%
  filter(!S_ID == 1114, 
         !S_ID == 1219, 
         !S_ID == 1023, 
         !S_ID == 2003, 
         !S_ID == 1015, 
         !S_ID == 1097) %>%  
  #select(S_ID:tc_race_other) %>%
  select(S_ID, TC_AgeMonths_Home_T30,PC_matage_home_T30,TC_Female,income2needs_T30,pc_ed) %>%
  #select(-c(box_wrist, inventorynotes_wrist, returned_wrist, returned_ankle, box_ankle, inventorynotes_ankle, Complete_Home_T30, GADays_Home_T30)) %>%
  rename("child_age_months_t30" = "TC_AgeMonths_Home_T30", 
         "mat_age_t30" = "PC_matage_home_T30","child_sex"="TC_Female","Maternal_Education"="pc_ed")



#Merge the parent and child characteristics with total number of hours worn
randomforest_data <- merge(attributes_weartimetotals, randomforest_demo_data, by="S_ID")%>%
  select(-c("parent_parentwrist_nighttotal","child_wrist_wktotal","child_ankle_wktotal","child_ankle_nighttotal","parent_parentwrist_wktotal", "S_ID")) %>%
  rename("PSI_Total_Parenting_Stress"="PSI_Mn_T30")%>%
  rename("PSI_Parenting_Stress_for_a_Difficult_Child"="PSI_DC_T30")%>%
  rename("MCDI_Overall_Language"="MCDI3_All_T30")%>%
  rename("ECBQ_Discomfort"="ECBQ_dis_T30")%>%
  rename("ECBQ_Effortful_Control"="ECBQ_EFFCO_T30")%>%
  rename("ECBQ_Negative_Affect"="ECBQ_NEGAF_T30")%>%
  rename("BIQ_Inhibition_to_Situational_Novelty"="BIQ_Nov_T30")%>%
  rename("BIQ_Total_Behavioral_Inhibition"="BIQ_Tot_T30")%>%
  rename("BIQ_Inhibition_to_Social_Situations"="BIQ_Soc_T30")%>%
  rename("PAS_Total_Anxiety"="PAS_AVG_T30")%>%
  rename("Parent_Wristband"="parent_parentwrist_daytotal")%>%
  rename("Child_Sex"="child_sex")%>%
  rename("Family_Income_to_Needs_Ratio"="income2needs_T30")%>%
  rename("Maternal_Education"="Maternal_Education")%>%
  rename("Maternal_Age"="mat_age_t30")%>%
  rename("Child_Age"="child_age_months_t30")%>%
  rename("ECBQ_Surgency"="ECBQ_SURGE_T30")





#Complete Cases
randomforest_data_complete_cases <- randomforest_data[complete.cases(randomforest_data), ]


ankle<-randomforest_data_complete_cases[,c(-12)]#Remove S_ID and wrist
wrist<-randomforest_data_complete_cases[,c(-13)] #Remove S_ID and ankle

# 5-fold cross validation
index = createFolds(wrist$child_wrist_daytotal, k = 5) 

# Seeing if about 20% of the records are in the testing set
kfold1 = index[[1]]
length(kfold1)/nrow(wrist)

# Creating vectors for parameters to be tuned
ntree_values = c(50, 250, 500) # number of decision trees 
p = dim(wrist)[2] - 1 # number of predictor variables in the dataset
mtry_values = c(sqrt(p), p/2, p) # number of predictors to be used in the model


# Creating an empty dataframe to save the confusion matrix metrics and variable importance
metrics = data.frame()
variable_importance_df = data.frame()

# Iterating through the cross validation folds
for (i in 1:length(index)){
  # Training data
  data_train = wrist[-index[[i]],]
  
  # Test data
  data_test = wrist[index[[i]],]
  
  
  # Creating empty lists and dataframes to store errors 
  reg_rf_pred_tune = list()
  rf <-randomForest(child_wrist_daytotal~.,data=wrist, ntree=500) 
  mtry <- tuneRF(wrist[-13],wrist$child_wrist_daytotal, ntreeTry=500,
                 stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
  best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
  
  
  # Now running RF on the entire training set with the tuned parameters
  # This will be done 5 times for each fold
  reg_rf <- randomForest(child_wrist_daytotal ~ ., data = data_train,
                         ntree = 500,
                         mtry = best.m)
  
  # Predicting on test set and adding the predicted values as an additional column to the test data
  data_test$wear_cat=case_when(data_test$child_wrist_daytotal<72~"L",
                               TRUE~"H")
  data_test$wear_cat=as.factor(data_test$wear_cat)
  data_test$Pred = predict(reg_rf, newdata = data_test, type = "response")
  data_test$Pred_cat=case_when(data_test$Pred<72~"L",
                               TRUE~"H")
  data_test$Pred_cat=as.factor(data_test$Pred_cat)
  levels(data_test$wear_cat) <- c("L", "H")
  levels(data_test$Pred_cat) <- c("L", "H")
  matrix = confusionMatrix(data = data_test$Pred_cat, 
                           reference = data_test$wear_cat)
  # Extracting accuracy, sens, spec, PPV, NPV and adding to the dataframe to take mean later
  matrix_values = data.frame(t(c(matrix$byClass[11])), t(c(matrix$byClass[1:4])))
  metrics = rbind(metrics, matrix_values)
  
  # Extracting variable importance
  variable_importance_values = data.frame(importance(reg_rf)) %>%
    rownames_to_column(var = "Predictor")
  variable_importance_df = rbind(variable_importance_df, variable_importance_values)
}

# Taking average across the 5 folds
metrics = metrics %>%
  summarise(`Balanced Accuracy` = mean(Balanced.Accuracy), Sensitivity = mean(Sensitivity), 
            Specificity = mean(Specificity), PPV = mean(Pos.Pred.Value), NPV = mean(Neg.Pred.Value,rm.na=TRUE))

variable_importance_df = variable_importance_df %>%
  group_by(Predictor) %>%
  summarise(MeanDecreaseGini = mean(IncNodePurity)) %>%
  # Sorting from highest to lowest
  arrange(-MeanDecreaseGini)

variable_importance_df
# Adding spaces between the variables that need the space
modified_variable_importance_df = variable_importance_df %>%
  mutate(Predictor = gsub("_", " ", Predictor))

# Saving the order of the variables from lowest to highest mean decrease gini by putting into a factor
predictor_order = rev(modified_variable_importance_df$Predictor)
modified_variable_importance_df$Predictor = factor(modified_variable_importance_df$Predictor, 
                                                   levels = predictor_order)

head(modified_variable_importance_df)
ggplot(data = modified_variable_importance_df , 
       aes(x = MeanDecreaseGini, y = Predictor, size = 2)) + 
  geom_point() + 
  theme_light() + 
  theme(axis.line = element_line(color = "black"), #making x and y axes black
        axis.text = element_text(size = 12), #changing size of x axis labels
        axis.title = element_text(face = "bold", size = rel(1.7)), #changes axis titles
        legend.title = element_text(face = 'bold', size = 14), #changes legend title
        legend.text = element_text(size = 12), #changes legend text
        strip.text.x = element_text(size = 15, face = "bold"), #changes size of facet x axis 
        strip.text.y = element_text(size = 15, face = "bold")) + #changes size of facet y axis 
  labs(x = 'Importance on Wrist Wear Time', y = ' Predictor Variable') + #changing axis labels 
  guides(size = "none")#removing size legend


################################################################################
index = createFolds(ankle$child_ankle_daytotal, k = 5) 

# Seeing if about 20% of the records are in the testing set
kfold1 = index[[1]]
length(kfold1)/nrow(ankle)

# Creating vectors for parameters to be tuned
ntree_values = c(50, 250, 500) # number of decision trees 
p = dim(ankle)[2] - 1 # number of predictor variables in the dataset
mtry_values = c(sqrt(p), p/2, p) # number of predictors to be used in the model


# Creating an empty dataframe to save the confusion matrix metrics and variable importance
metrics = data.frame()
variable_importance_df = data.frame()

# Iterating through the cross validation folds
for (i in 1:length(index)){
  # Training data
  data_train = ankle[-index[[i]],]
  
  # Test data
  data_test = ankle[index[[i]],]
  
  #data_test$child_ankle_wktotal <- factor(data_test$child_ankle_wktotal)
  
  
  # Creating empty lists and dataframes to store errors 
  reg_rf_pred_tune = list()
  rf <-randomForest(child_ankle_daytotal~.,data=ankle, ntree=500) 
  mtry <- tuneRF(ankle[-13],ankle$child_ankle_daytotal, ntreeTry=500,
                 stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
  best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
  
  # Tuning parameters: using ntree and mtry values to determine which combination yields the smallest OOB error 
  # from the validation datasets
  
  # Now running RF on the entire training set with the tuned parameters
  # This will be done 5 times for each fold
  reg_rf <- randomForest(child_ankle_daytotal ~ ., data = data_train,
                         ntree = 500,
                         mtry = best.m)
  
  # Predicting on test set and adding the predicted values as an additional column to the test data
  data_test$wear_cat=case_when(data_test$child_ankle_daytotal<72~"L",
                               TRUE~"H")
  data_test$wear_cat=as.factor(data_test$wear_cat)
  data_test$Pred = predict(reg_rf, newdata = data_test, type = "response")
  data_test$Pred_cat=case_when(data_test$Pred<72~"L",
                               TRUE~"H")
  data_test$Pred_cat=as.factor(data_test$Pred_cat)
  levels(data_test$wear_cat) <- c("L", "H")
  levels(data_test$Pred_cat) <- c("L", "H")
  matrix = confusionMatrix(data = data_test$Pred_cat, 
                           reference = data_test$wear_cat)
  # Extracting accuracy, sens, spec, PPV, NPV and adding to the dataframe to take mean later
  matrix_values = data.frame(t(c(matrix$byClass[11])), t(c(matrix$byClass[1:4])))
  metrics = rbind(metrics, matrix_values)
  
  # Extracting variable importance
  variable_importance_values = data.frame(importance(reg_rf)) %>%
    rownames_to_column(var = "Predictor")
  variable_importance_df = rbind(variable_importance_df, variable_importance_values)
}

# Taking average across the 5 folds
metrics = metrics %>%
  summarise(`Balanced Accuracy` = mean(Balanced.Accuracy), Sensitivity = mean(Sensitivity), 
            Specificity = mean(Specificity), PPV = mean(Pos.Pred.Value), NPV = mean(Neg.Pred.Value))

#Variable importance is a reflection of how much the performance of the model decreases when the variable is removed
variable_importance_df = variable_importance_df %>%
  group_by(Predictor) %>%
  summarise(MeanDecreaseGini = mean(IncNodePurity)) %>%
  #summarise(MeanDecreaseGini_std = stdev(IncNodePurity)) %>%
  # Sorting from highest to lowest
  arrange(-MeanDecreaseGini)

variable_importance_df
# Adding spaces between the variables that need the space
modified_variable_importance_df = variable_importance_df %>%
  mutate(Predictor = gsub("_", " ", Predictor))

# Saving the order of the variables from lowest to highest mean decrease gini by putting into a factor
predictor_order = rev(modified_variable_importance_df$Predictor)
modified_variable_importance_df$Predictor = factor(modified_variable_importance_df$Predictor, 
                                                   levels = predictor_order)

head(modified_variable_importance_df)

ggplot(data = modified_variable_importance_df , 
       aes(x = MeanDecreaseGini, y = Predictor, size = 2)) + 
  geom_point() + 
  
  theme_light() + 
  theme(axis.line = element_line(color = "black"), #making x and y axes black
        axis.text = element_text(size = 12), #changing size of x axis labels
        axis.title = element_text(face = "bold", size = rel(1.7)), #changes axis titles
        legend.title = element_text(face = 'bold', size = 14), #changes legend title
        legend.text = element_text(size = 12), #changes legend text
        strip.text.x = element_text(size = 15, face = "bold"), #changes size of facet x axis 
        strip.text.y = element_text(size = 15, face = "bold")) + #changes size of facet y axis 
  labs(x = 'Importance on Ankle Wear Time', y = ' Predictor Variable') + #changing axis labels 
  
  guides(size = "none")#removing size legend


```

























maunscript plots

FIGURE 1-- box plot of child wrist and ankle band wear time by day and night time 
```{r}

figure1 <- ggplot(wksumsep, aes(x=day, y=hours, fill=position)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.2)) +
  stat_compare_means(method = "wilcox.test",
                     paired= TRUE,
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("Child Ankle" = "lavender", 
                               "Child Wrist" = "#9999CC")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Total Hours") +
  xlab("Time of Day")
  
figure1

jpeg("2_OutputGraphs/raw/figure1.jpeg",
    width = 7, height = 5, units="in", res=1200)
figure1
invisible(dev.off())

```
FIGURE 2 was generated in powerpoint and illustrator

FIGURE 3 
boxplots with time of day (x) by total wear hours (y) 
including paired wilcox comparing wrist (dark) by ankle (light)
```{r}
# change the name of the 
demo_weartime_sep <- demo_weartime_sep %>% 
  mutate(position2=case_when(position == "ankle" ~ "Child Ankle", 
                            position == "wrist" ~ "Child Wrist"))


# paired, stratified by sex
figure3 <- ggplot(demo_weartime_sep, aes(x=sex, y=hours, fill= position2)) + 
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.2)) +
  stat_compare_means(method = "wilcox.test",
                     paired= TRUE,
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("Child Ankle" = "lavender", 
                               "Child Wrist" = "#9999CC")) +
  scale_x_discrete(labels = c("Female", "Male")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Total Daytime Hours") +
  xlab("Sex")
  
figure3



jpeg("2_OutputGraphs/raw/figure3.jpeg",
    width = 7, height = 5, units="in", res=1200)
figure3
invisible(dev.off())

```

FIGURE 4
select attributes for wristband
boxplots with categorized hours (x) by attribute scale (y) and wilcox test
scatter plots with total daytime hours (x) and attribute scale (y) and spearman correlation
attributes included: PAS and MCDI
```{r}

#separate wrist plots to be combined with patchwork
long_wk_child_day_wrist_pas <- long_wk_child_day %>% 
  filter(attribute== "PAS_AVG_T30", 
         position== "wrist")

#boxplot comparing categorized weartime (wilcox rank sum test)
figure4_pas <- ggplot(long_wk_child_day_wrist_pas, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#f8c198", 
                               "72 or more hours" = "#e26f1a")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "PAS: Average\nTwo-Group Comparison",
       subtitle= "(p = 0.038)") 

figure4_pas

#jpeg("2_OutputGraphs/raw/figure4_pas.jpeg",
#    width = 160, height = 300, res=1000)
#figure4_pas
#invisible(dev.off())




#scatter with spearman correlation
figure4_pas_scat <- ggplot(long_wk_child_day_wrist_pas, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#e26f1a", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "PAS: Average\nCorrelation",
    subtitle= "(rho = 0.19, p = 0.085)")

figure4_pas_scat


#jpeg("2_OutputGraphs/raw/figure4_pas_scat.jpeg",
#    width = 2000, height = 2000)
#figure4_pas_scat
#invisible(dev.off())



#mcdi
long_wk_child_day_wrist_mcdi <- long_wk_child_day %>% 
  filter(attribute== "MCDI3_All_T30", 
         position== "wrist")


figure4_mcdi <- ggplot(long_wk_child_day_wrist_mcdi, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#acdaa6", 
                               "72 or more hours" = "#387830")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "MCDI: Overall Language\nTwo-Group Comparison",
       subtitle= "(p = 0.015)") 

figure4_mcdi

#jpeg("2_OutputGraphs/raw/figure4_mcdi.jpeg",
#    width = 1600, height = 2800)
#figure4_mcdi
#invisible(dev.off())




#scatter
figure4_mcdi_scat <- ggplot(long_wk_child_day_wrist_mcdi, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#387830", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "MCDI: Overall Language\nCorrelation",
    subtitle= "(rho = 0.26, p = 0.010)")

figure4_mcdi_scat


#jpeg("2_OutputGraphs/raw/figure4_mcdi_scat.jpeg",
#    width = 2000, height = 2000)
#figure4_mcdi_scat
#invisible(dev.off())



#combine the FOUR plots to make final figure 
figure4 <-  figure4_mcdi + figure4_pas + figure4_mcdi_scat + figure4_pas_scat + 
  plot_layout(ncol=2, heights = c(2, 1)) + plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  theme(plot.tag = element_text(size = 18, face = "bold", ))


figure4

jpeg("2_OutputGraphs/figure4.jpeg",
    width = 9, height = 9, units="in", res=1200)
figure4
invisible(dev.off())



```

FIGURE 5
select attributes for ankle band
boxplots with categorized hours (x) by attribute scale (y) and wilcox test
scatter plots with total daytime hours (x) and attribute scale (y) and spearman correlation
attributes included: PAS and MCDI
```{r}

#figure 4 individual plots (ankle)

#set up 
long_wk_child_day_ankle_biq_nov <- long_wk_child_day %>% 
  filter(attribute== "BIQ_Nov_T30", 
         position== "ankle")

long_wk_child_day_ankle_biq_soc <- long_wk_child_day %>% 
  filter(attribute== "BIQ_Soc_T30", 
         position== "ankle")

long_wk_child_day_ankle_biq_tot <- long_wk_child_day %>% 
  filter(attribute== "BIQ_Tot_T30", 
         position== "ankle")

long_wk_child_day_ankle_ecbq_effco <- long_wk_child_day %>% 
  filter(attribute== "ECBQ_EFFCO_T30", 
         position== "ankle")


## FIRST COLUMN ##

#ankle: biq novelty 
figure5_biq_nov <- ggplot(long_wk_child_day_ankle_biq_nov, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#acdaa6", 
                               "72 or more hours" = "#387830")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "BIQ: Novelty\nTwo-Group Comparison",
       subtitle= "(p = 0.003)")
  
  
figure5_biq_nov


#scatter
figure5_biq_nov_scat <- ggplot(long_wk_child_day_ankle_biq_nov, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#387830", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "BIQ: Novelty\nCorrelation",
    subtitle= "(rho = 0.31, p = 0.002)")

figure5_biq_nov_scat













## COLUMN 2 ##

#ankle: biq social 
figure5_biq_soc <- ggplot(long_wk_child_day_ankle_biq_soc, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#f8c198", 
                               "72 or more hours" = "#e26f1a")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "BIQ: Social\nTwo-Group Comparison",
       subtitle= "(p = 0.001)")

figure5_biq_soc



figure5_biq_soc_scat <- ggplot(long_wk_child_day_ankle_biq_soc, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#e26f1a", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "BIQ: Social\nCorrelation",
    subtitle= "(rho = 0.35, p < 0.001)")
  
  
  
figure5_biq_soc_scat











# COLUMN 3 ##


#ankle: biq total 
figure5_biq_tot <- ggplot(long_wk_child_day_ankle_biq_tot, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#dbabf0", 
                               "72 or more hours" = "#af69cd")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "BIQ: Total\nTwo-Group Comparison",
       subtitle= "(p = 0.001)")
  
figure5_biq_tot



#scatter
figure5_biq_tot_scat <- ggplot(long_wk_child_day_ankle_biq_tot, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#af69cd", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.1, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "BIQ: Total\nCorrelation",
    subtitle= "(rho = 0.34, p < 0.001)")

figure5_biq_tot_scat




# COLUMN 4 ##

#ankle: ecbq effco
figure5_ecbq_effco <- ggplot(long_wk_child_day_ankle_ecbq_effco, aes(x=cat_hours, y=value, fill=cat_hours)) +
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1, position= position_dodge(width=5)) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.3)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#e997ae", 
                               "72 or more hours" = "#c84268")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Categorized Hours") +
  labs(title= "ECBQ: Effortful Control\nTwo-Group Comparison",
       subtitle= "(p = 0.016)")
  
figure5_ecbq_effco





figure5_ecbq_effco_scat <- ggplot(long_wk_child_day_ankle_ecbq_effco, aes(x=hours, y=value)) + 
  geom_point(size= 1.5) + 
  geom_smooth(method="lm", se=F, size=1.5, color="#c84268", linetype="dashed") +
  scale_shape_manual(values= c(16)) +
  #stat_cor(method = "spearman", label.x=c(3,5), label.y=c(0.7,0.75), size=10, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Attribute Scale") +
  xlab("Total Daytime Hours") + 
  labs(title= "ECBQ: Effortful Control\nCorrelation",
    subtitle= "(rho = -0.2, p = 0.041)")


figure5_ecbq_effco_scat



#figures combined

#combine the FOUR plots to make final figure 
figure5 <-  figure5_biq_nov + figure5_biq_soc + figure5_biq_tot + figure5_ecbq_effco + figure5_biq_nov_scat + figure5_biq_soc_scat + figure5_biq_tot_scat + figure5_ecbq_effco_scat + plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(ncol=4, heights = c(2, 1)) &
  theme(plot.tag = element_text(size = 18, face = "bold"))
  
figure5

jpeg("2_OutputGraphs/figure5.jpeg",
    width = 16, height = 9, units="in", res=1200)
figure5
invisible(dev.off())


```



boxplot of age (supp figure 1)
```{r}

# wrist boxplot with wear time category (x) by age (y)
figure_age_wrist <- ggplot(demo_weartime_wrist, aes(x=cat6day, y=age, fill=cat6day)) + 
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.2)) +
  stat_compare_means(method = "wilcox.test",
                     # Changing the value to asterisks and moving to the middle of the plot
                     label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#A6CAED", 
                               "72 or more hours" = "#6699CC")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Child Age (months)") +
  xlab("Categorized Hours")
  

figure_age_wrist

wilcox_age_wrist <- wilcox.test(age ~ cat6day, data= demo_weartime_wrist)
formatted_p_value <- format.pval(wilcox_age_wrist$p.value, digits = 3)
print(formatted_p_value)

# ankle boxplot with wear time category (x) by age (y)

figure_age_ankle <- ggplot(demo_weartime_ankle, aes(x=cat6day, y=age, fill=cat6day)) + 
  geom_boxplot(outlier.shape = NA, lwd=1, alpha=1) +
  geom_jitter(size=1.5, position = position_jitterdodge(0.2)) +
  #stat_compare_means(method = "wilcox.test",                                                       ## removed to not have the "ns" showing
                     # Changing the value to asterisks and moving to the middle of the plot
                     #label = "p.signif", label.x.npc = "center", size = 12) +
  scale_fill_manual(values = c("< 72 hours" = "#acdaa6", 
                               "72 or more hours" = "#387830")) +
  scale_x_discrete(labels = c("<72 Hours", "≥72 Hours")) +
  # Adding padding y axis 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_classic() +
  theme(axis.text=element_text(size=14, color = "black"), 
        axis.text.x=element_text(vjust=-1),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 7, b=0, l=0)),
        
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(),
        
        axis.title=element_text(size=16, face="bold", color = "black"),
        axis.title.x=element_text(vjust=-3), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 10, b=0, l=0)),
        
        title=element_text(size=14),
        plot.title=element_text(vjust=5, hjust=0.5, face="bold"),
        #plot.title.position="plot",
        subtitle=element_markdown(),
        plot.subtitle=element_text(vjust=1, hjust=1, size=14), 
        
        plot.margin= unit(c(0.3, 0.3, 0.3, 0.3), "inches"), 
        panel.border = element_rect(color = "black", fill=NA, linewidth=1.5),
        legend.position= "none") +
  ylab("Child Age (months)") +
  xlab("Categorized Hours")
  

figure_age_ankle


supp_figure1 <-  figure_age_wrist + figure_age_ankle + plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 18, face = "bold"))

supp_figure1

jpeg("2_OutputGraphs/supp_figure1.jpeg",
    width = 10, height = 5, units="in", res=1200)
supp_figure1
invisible(dev.off())

```




cleaned figure, not in manuscript
scatter plots of female and male wrist (x) by ankle (y) weartime 
```{r}

#individual plots for figure 

# female
scat_f <- ggplot(demo_weartime_female, aes(x=wrist_hours, y=ankle_hours, shape=sex)) + 
  geom_point(size= 24) + 
  geom_smooth(method="lm", se=F, size=7)+
  scale_shape_manual(values= c(15)) +
  #labs(title= "Figure 2") +
  ylab("Ankle Daytime Hours") +
  xlab("Wrist Daytime Hours") + 
  labs(title= "(A) Female") +
  stat_cor(method = "spearman", label.x=c(3,5), label.y=c(30,35), size=40, fontface="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=100, face="bold", color = "black"), 
        axis.title=element_text(size=110, face="bold", color = "black"),
        axis.text.x=element_text(vjust=-1),
        axis.title.x=element_text(vjust=-1.5), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 50, b=0, l=0)),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 30, b=0, l=0)),
        axis.line=element_line((size=10), color="black"),
        title=element_text(size=100, face="bold"),
        plot.title=element_text(vjust=2), 
        legend.title=element_text(size=100),
        legend.text=element_text(size=100), 
        plot.margin= unit(c(1.5, 1.5, 1.5, 1.5), "inches"), 
        legend.position= "none",
        panel.border = element_rect(color = "black", fill=NA, linewidth=10))

scat_f

#jpeg("2_OutputGraphs/raw/scat_f.jpeg",
#    width = 2400, height = 2400)
#scat_f
#invisible(dev.off())



# male
scat_m <- ggplot(demo_weartime_male, aes(x=wrist_hours, y=ankle_hours, shape=sex)) + 
  geom_point(size= 24) + 
  geom_smooth(method="lm", se=F, size=7)+
  scale_shape_manual(values= c(16)) +
  #labs(title= "Figure 2") +
  ylab("Ankle Daytime Hours") +
  xlab("Wrist Daytime Hours") + 
  labs(title= "(B) Male") +
  stat_cor(method = "spearman", label.x=c(3,5), label.y=c(30,35), size=40, face="bold") + 
  theme_classic() +
  theme(axis.text=element_text(size=100, face="bold", color = "black"), 
        axis.title=element_text(size=110, face="bold", color = "black"),
        axis.text.x=element_text(vjust=-1),
        axis.title.x=element_text(vjust=-1.5), 
        axis.title.y=element_text(margin= ggplot2::margin(t=0, r= 50, b=0, l=0)),
        axis.text.y=element_text(margin= ggplot2::margin(t=0, r= 30, b=0, l=0)),
        axis.line=element_line((size=10), color="black"),
        title=element_text(size=100, face="bold"),
        plot.title=element_text(vjust=2), 
        legend.title=element_text(size=100),
        legend.text=element_text(size=100), 
        plot.margin= unit(c(1.5, 1.5, 1.5, 1.5), "inches"), 
        legend.position= "none",
        panel.border = element_rect(color = "black", fill=NA, linewidth=10))

scat_m

#jpeg("2_OutputGraphs/raw/scat_m.jpeg",
#    width = 2400, height = 2400)
#scat_m
#invisible(dev.off())

#combine the two plots to make final figure 
scatter_bysex <- scat_f + scat_m +
  #plot_layout(widths = c(3, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 18, face = "bold"))
  
scatter_bysex


```



package citations to add to endnote for the manuscript
```{r}

citation("openxlsx") # for reading and writing out files
citation("table1") # for making tables
citation("tidyverse") # for data organization and manipulation
citation("ggpubr") 
citation("rstatix")
citation("ltm") # cronbach's alpha
citation("FSA") # dunns test
citation("patchwork")

```

